### 2023-07-10 1.introduction.tex
Update abstract

% !TeX spellcheck = en_US
\section{Introduction}
\label{sec_introduction}

### 2023-07-10 paper.tex
Update abstract

\documentclass[pageno]{jpaper}

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

\usepackage[normalem]{ulem}

\begin{document}

\title{
Static Race Detection in OS Kernels by Mining Locking Rules}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}

To take advantage of multiple-core architecture of modern CPUs, an operating 
system (OS) is often designed to be highly concurrent, and processes shared 
data structure in parallel. To ensure safe access to shared data structures, 
fine-grained locking mechanisms are used to avoid data races among different 
threads. However, it is hard to determine whether a lock should be used for a 
specific data structure field (locking rules) even for an expert developer, due 
to poor documentation and complicated logic of OS code. As a result, OS kernel 
is prone to data races, because necessary locks may be missed by mistake. 
Static analysis is a common technique to help improve code quality, but it is 
quite challenging to detect data races in OS kernels automatically, because of 
lack of knowledge of locking rules and high complexity of concurrent execution.

In this paper, we design a practical static analysis approach named RaceMiner, 
to effectively detect harmful races that can trigger memory or logical bugs in 
OS kernels by mining locking rules. RaceMiner first employs an alias-aware rule 
mining method to automatically deduce locking rules, and detects data races 
caused by violation of these rules. And then performs a lock-usage analysis to 
filter out false positives caused by concurrency. At last, RaceMiner extracts 
harmful data races from all detected data races through pattern-based 
estimation. We have evaluated RaceMiner on Linux 6.2, and find xxx data races, 
with a false positive rate of xxx. Among these data races, xxx are estimated to 
be harmful. We have reported these harmful bugs to Linux kernel developers, and 
xxx of them have been confirmed.

\end{abstract}

\input{1.introduction}

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2023-07-14 1.introduction.tex
Update Instruction

@@
 \label{sec_introduction}
+
+To take advantage of multiple-core architecture of modern CPUs, an OS kernel is 
+often designed to be highly concurrent. However, concurrent execution can 
+inevitably introduce concurrency bugs. Data race is a common kind of 
+concurrency bugs that occurs when multiple threads access the same variable 
+concurrently and at least one of the accesses is a write. Some data races are 
+quite dangerous and can cause serious bugs such as null-pointer dereferences, 
+data inconsistency and infinite loops.
+
+To detect data races effectively, some approaches~\cite{Boyapati:OOPSLA02, 
+Anderson:PLDI08, Anderson:PLDI09, Zhou:MICRO19, Flanagan:PASTE01, 
+Flanagan:PLDI00, Sadowski:PLATEAU14, ClangThreadSafety, Blackshear:OOPSLA18} 
+rely on developers to 
+supply annotations that describe the locking discipline such as which lock 
+protects a specific data structure field, and detect data races by exploiting 
+data structure field accesses that violate these disciplines. For example, 
+Clang thread safety analysis~\cite{ClangThreadSafety} requires developers to 
+label a variable and the lock to protect it with a {\tt GUARDED\_BY} attribute. 
+And then detects data races by finding variable accesses that violate the given 
+attribute. However, such annotation-based approaches are not suitable for race 
+detection in OS kernels, because even an expert developer can not provide 
+accurate annotations, due to poor documentation and complicated logic of OS 
+codes.
+
+Some static approaches~\cite{Choi:PLDI02, Engler:SOSP03, Voung:FSE07, 
+Pratikakis:PLDI06, Naik:PLDI06} employ lockset-based analysis to detect data 
+races automatically. They first collect locks that are acquired when a given 
+variable is accessed, and then mine locking rules about whether a given 
+variable should be protected by a specific lock. At last, they detect data 
+races by checking whether variable accesses violate the mined locking rules. 
+However, they do not consider alias relations~\cite{Voung:FSE07, Engler:SOSP03} 
+or just use imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, 
+Pratikakis:PLDI06, Naik:PLDI06}, and thus can introduce both false positives 
+and false negatives.
+
+Dynamic analysis can acquire run-time information such as memory address, and 
+thus complex alias analysis can be avoided. Therefore, some 
+approaches~\cite{Lochmann:EuroSys19, Lu:SOSP07, Lu:FSE18, Joshi:ASE08, 
+Liu:NSDI07} detect data races dynamically to improve precision. They mine 
+implicit code rules in softwares by statistically analyzing execution traces, 
+and then use the mined rules to detect related bugs. For example, 
+LockDoc~\cite{Lochmann:EuroSys19} records accesses to data structure fields and 
+lock acquisitions of an instrumented Linux kernel, and then infers locking 
+rules from the recorded accesses. After that, LockDoc automatically locates 
+data structure fields access that violate the inferred locking rules to detect 
+concurrency bugs such as data races. However, dynamic analysis suffers from low 
+code coverage, and thus the locking rules inferred through execution traces can 
+be imprecise. Besides, dynamic analyses are hard to deploy, because they need 
+to run OS kernels with instruments. At last, it is difficult for dynamic 
+tools to estimate the harmfulness of a data race, because race conditions are 
+hard to trigger by running the checked software~\cite{Fonseca:DSN10, 
+Burckhardt:ASPLOS10, Liu:FSE14, Zhou:EASE15}.
+
+In this paper, we design a practical static analysis approach named RaceMiner, 
+to automatically detect data races and estimate the harmfulness of these data 
+races in OS kernels effectively. RaceMiner consists of three key techniques:
+
+(T1) RaceMiner uses an {\em alias-aware rule mining method} to automatically 
+deduce locking rules. We observe that an access to a data structure field is 
+often protected by the lock in the same data structure as the accessed field. 
+And this relationship between accessed field and the protecting lock can be 
+effectively inferred from a special data structure named alias 
+graph~\cite{Li:ASPLOS22, Kastrinis:CC18}. Specifically, fields in the same data 
+structure have the same ancestor in the alias graph, and thus whether an 
+accessed field and a specific lock are in the same data structure can be 
+inferred by finding a common ancestor in the alias graph. Moreover, with 
+benefits from precise field-sensitive alias relationships of alias graph, 
+RaceMiner can accurately extract the accessed field and its corresponding 
+protecting lock. After obtaining the accessed field and protecting lock pairs, 
+RaceMiner calculates the proportion of field accesses protected by a specific 
+lock in all field accesses. If the proportion is larger than a given threshold, 
+the accessed field is determined to be protected by the protecting lock.
+
+(T2) RaceMiner uses a {\em lock-usage analysis} to filter out false positives 
+caused by concurrency. After mining locking rules, RaceMinder detects data 
+races by checking whether a given data structure field access violates the 
+locking rules. To reduce false negatives, RaceMinder conservatively assumes 
+every two functions can be executed concurrently, and thus can introduce false 
+positives. However, we observe that each kernel module has an initialization 
+phase, after which many functions can be called concurrently by other modules. 
+When performing initialization, the kernel module serially initialized the lock 
+and prepare other data for subsequent operations. And thus functions with the 
+lock initialization and functions called by them tend not to be executed 
+concurrently. Based on this observation, RaceMiner extracts all functions 
+reachable from the function with a lock initialization operation such as {\tt 
+spin\_lock\_init()}, and suppose that this function can not be executed in 
+parallel.
+
+(T3) RaceMiner uses a {\em pattern-based estimation} to extract harmful races 
+that can trigger memory or logical bugs such as null-pointer dereferences and 
+data inconsistency. Some data races are benign, and developers do not put 
+effort into repairing them because they can not cause serious memory or logical 
+bugs. Therefore, it is important to estimate the harm a data race can bring. We 
+observe that harmful data races can be estimated by given patterns. For 
+example, if a raced data performs as an operand of a dereference instruction, a 
+null-pointer dereference can occur; if a raced data is accessed for more than 
+once and data inconsistency can occur. Based on this observation, RaceMiner 
+uses some patterns to extract harmful races that can cause memory or logical 
+bugs automatically.
+
+We have implemented RaceMiner with LLVM~\cite{clang} and Z3~\cite{z3}. 
+RaceMiner performs automated static analysis on the LLVM bytecode of the 
+checked OS kernel. Overall, we make three main contributions in this paper:
+
+\begin{itemize}
+	\item We analyze the challenges of data race detection in OS kernels, and 
+	propose three key techniques to address	these challenges: (T1) an {\em 
+	alias-aware rule mining method} to automatically deduce locking rules; (T2) 
+	a {\em lock-usage analysis} to filter out false positives caused by 
+	concurrency; (T3) a {\em pattern-based estimation} to extract harmful data 
+	races that can trigger memory or logical bugs such as null-pointer 
+	dereferences and data inconsistency.
+	\item Based on these three key techniques, we design a practical static 
+	analysis approach named RaceMiner, to effectively detect data races and 
+	estimate the harmfulness of these data races in OS kernels.
+	\item We have evaluated RaceMiner on Linux 6.2, and find xxx data races, 	
+	with a false positive rate of xxx. Among these data races, xxx are 
+	estimated to be harmful. We have reported these harmful bugs to Linux 
+	kernel developers, and xxx of them have been confirmed.
+\end{itemize}
+
+The rest of this paper is organized as follows. Section~\ref{sec_motivation} 
+introduces the motivation and challenges of data race detection in OS kernels. 
+Section~\ref{sec_technique} introduces our key techniques to address these 
+challenges. Section~\ref{sec_framework} introduces RaceMiner. 
+Section~\ref{sec_evaluation} shows our evaluation. Section~\ref{sec_discussion} 
+makes a discussion about RaceMiner. Section~\ref{sec_related} presents related 
+work, and Section~\ref{sec_conclusion} concludes this paper.

### 2023-07-14 2.motivation.tex
Update Instruction

% !TeX spellcheck = en_US
\section{Motivation}
\label{sec_motivation}

### 2023-07-14 3.technique.tex
Update Instruction

% !TeX spellcheck = en_US
\section{Race Detection by Mining Locking Rules}
\label{sec_technique}

### 2023-07-14 4.framework.tex
Update Instruction

% !TeX spellcheck = en_US
\section{Framework}
\label{sec_framework}

### 2023-07-14 5.evaluation.tex
Update Instruction

% !TeX spellcheck = en_US
\section{Evaluation}
\label{sec_evaluation}

### 2023-07-14 6.discussion.tex
Update Instruction

% !TeX spellcheck = en_US
\section{Discussion}
\label{sec_discussion}

### 2023-07-14 7.related.tex
Update Instruction

% !TeX spellcheck = en_US
\section{Related Work}
\label{sec_related}

### 2023-07-14 8.conclusion.tex
Update Instruction

% !TeX spellcheck = en_US
\section{Conclusion}
\label{sec_conclusion}

### 2023-07-14 paper.tex
Update Instruction

@@
 \usepackage[normalem]{ulem}
+
+\usepackage[numbers, sort]{natbib}
 
@@
 system (OS) is often designed to be highly concurrent, and processes shared 
-data structure in parallel. To ensure safe access to shared data structures, 
+data structures in parallel. To ensure safe access to shared data structures, 
 fine-grained locking mechanisms are used to avoid data races among different 
@@
 filter out false positives caused by concurrency. At last, RaceMiner extracts 
-harmful data races from all detected data races through pattern-based 
+harmful data races from all detected data races through a pattern-based 
 estimation. We have evaluated RaceMiner on Linux 6.2, and find xxx data races, 
@@
 \input{1.introduction}
+\input{2.motivation}
+\input{3.technique}
+\input{4.framework}
+\input{5.evaluation}
+\input{6.discussion}
+\input{7.related}
+\input{8.conclusion}
 
+\footnotesize
 \bibliographystyle{plain}

### 2023-07-17 paper.tex
Update paper.tex

@@
 \usepackage[normalem]{ulem}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{xspace}
 
 \usepackage[numbers, sort]{natbib}
+
+\newcommand{\sys}{\mbox{RaceMiner}\xspace}
 
@@
 To take advantage of multiple-core architecture of modern CPUs, an operating 
-system (OS) is often designed to be highly concurrent, and processes shared 
-data structures in parallel. To ensure safe access to shared data structures, 
-fine-grained locking mechanisms are used to avoid data races among different 
-threads. However, it is hard to determine whether a lock should be used for a 
-specific data structure field (locking rules) even for an expert developer, due 
-to poor documentation and complicated logic of OS code. As a result, OS kernel 
-is prone to data races, because necessary locks may be missed by mistake. 
-Static analysis is a common technique to help improve code quality, but it is 
-quite challenging to detect data races in OS kernels automatically, because of 
-lack of knowledge of locking rules and high complexity of concurrent execution.
+system (OS) is expected to handle concurrency, inevitably introducing many 
+concurrent issues. And data race is one of the most harmful concurrent issues 
+that can trigger memory or logical bugs. A data race is caused by wrongly 
+accessing shared data in different threads. To ensure correct access to shared 
+data, fine-grained locking mechanisms are introduced. However, it is hard to 
+determine whether a lock should be used for a specific data structure field 
+(locking rules) even for an expert developer, due to poor documentation and 
+complicated logic of OS code. As a result, OS kernel is prone to data races, 
+because necessary locks may be missed by mistake. Static analysis is a common 
+technique to help improve code quality, but it is quite challenging to detect 
+data races in OS kernels automatically, because of lack of knowledge of locking 
+rules and high complexity of concurrent execution.
 
-In this paper, we design a practical static analysis approach named RaceMiner, 
-to effectively detect harmful races that can trigger memory or logical bugs in 
-OS kernels by mining locking rules. RaceMiner first employs an alias-aware rule 
-mining method to automatically deduce locking rules, and detects data races 
-caused by violation of these rules. And then performs a lock-usage analysis to 
-filter out false positives caused by concurrency. At last, RaceMiner extracts 
-harmful data races from all detected data races through a pattern-based 
-estimation. We have evaluated RaceMiner on Linux 6.2, and find xxx data races, 
-with a false positive rate of xxx. Among these data races, xxx are estimated to 
-be harmful. We have reported these harmful bugs to Linux kernel developers, and 
-xxx of them have been confirmed.
+In this paper, we design a practical static analysis approach named \sys, to 
+effectively detect harmful races that can trigger memory or logical bugs in OS 
+kernels by mining locking rules. \sys first employs an alias-aware rule mining 
+method to automatically deduce locking rules, and detects data races caused by 
+violation of these rules. And then performs a lock-usage analysis to filter out 
+false positives caused by concurrency. At last, \sys extracts harmful data 
+races from all detected data races through a pattern-based estimation. We have 
+evaluated \sys on Linux 6.2, and find xxx data races, with a false positive 
+rate of xxx. Among these data races, xxx are estimated to be harmful. We have 
+reported these harmful bugs to Linux kernel developers, and xxx of them have 
+been confirmed.
 

### 2023-07-19 1.introduction.tex
Add motivation

@@
 To take advantage of multiple-core architecture of modern CPUs, an OS kernel is 
-often designed to be highly concurrent. However, concurrent execution can 
-inevitably introduce concurrency bugs. Data race is a common kind of 
-concurrency bugs that occurs when multiple threads access the same variable 
-concurrently and at least one of the accesses is a write. Some data races are 
-quite dangerous and can cause serious bugs such as null-pointer dereferences, 
-data inconsistency and infinite loops.
+expected to handle concurrency. However, concurrent execution can inevitably 
+introduce concurrency issues. Data race is one of the most harmful concurrency 
+issues that occurs when multiple threads access the same variable concurrently 
+and at least one of the accesses is a write. Some data races are quite 
+dangerous and can cause serious bugs such as null-pointer dereferences, data 
+inconsistency and infinite loops.
 
@@
 
-In this paper, we design a practical static analysis approach named RaceMiner, 
+In this paper, we design a practical static analysis approach named \sys, 
 to automatically detect data races and estimate the harmfulness of these data 
-races in OS kernels effectively. RaceMiner consists of three key techniques:
+races in OS kernels effectively. \sys consists of three key techniques:
 
-(T1) RaceMiner uses an {\em alias-aware rule mining method} to automatically 
+(T1) \sys uses an {\em alias-aware rule mining method} to automatically 
 deduce locking rules. We observe that an access to a data structure field is 
@@
 benefits from precise field-sensitive alias relationships of alias graph, 
-RaceMiner can accurately extract the accessed field and its corresponding 
+\sys can accurately extract the accessed field and its corresponding 
 protecting lock. After obtaining the accessed field and protecting lock pairs, 
-RaceMiner calculates the proportion of field accesses protected by a specific 
+\sys calculates the proportion of field accesses protected by a specific 
 lock in all field accesses. If the proportion is larger than a given threshold, 
@@
 
-(T2) RaceMiner uses a {\em lock-usage analysis} to filter out false positives 
-caused by concurrency. After mining locking rules, RaceMinder detects data 
+(T2) \sys uses a {\em lock-usage analysis} to filter out false positives 
+caused by concurrency. After mining locking rules, \sys detects data 
 races by checking whether a given data structure field access violates the 
-locking rules. To reduce false negatives, RaceMinder conservatively assumes 
+locking rules. To reduce false negatives, \sys conservatively assumes 
 every two functions can be executed concurrently, and thus can introduce false 
@@
 lock initialization and functions called by them tend not to be executed 
-concurrently. Based on this observation, RaceMiner extracts all functions 
+concurrently. Based on this observation, \sys extracts all functions 
 reachable from the function with a lock initialization operation such as {\tt 
@@
 
-(T3) RaceMiner uses a {\em pattern-based estimation} to extract harmful races 
+(T3) \sys uses a {\em pattern-based estimation} to extract harmful races 
 that can trigger memory or logical bugs such as null-pointer dereferences and 
@@
 null-pointer dereference can occur; if a raced data is accessed for more than 
-once and data inconsistency can occur. Based on this observation, RaceMiner 
+once and data inconsistency can occur. Based on this observation, \sys 
 uses some patterns to extract harmful races that can cause memory or logical 
@@
 
-We have implemented RaceMiner with LLVM~\cite{clang} and Z3~\cite{z3}. 
-RaceMiner performs automated static analysis on the LLVM bytecode of the 
+We have implemented \sys with LLVM~\cite{clang} and Z3~\cite{z3}. 
+\sys performs automated static analysis on the LLVM bytecode of the 
 checked OS kernel. Overall, we make three main contributions in this paper:
@@
 	\item Based on these three key techniques, we design a practical static 
-	analysis approach named RaceMiner, to effectively detect data races and 
+	analysis approach named \sys, to effectively detect data races and 
 	estimate the harmfulness of these data races in OS kernels.
-	\item We have evaluated RaceMiner on Linux 6.2, and find xxx data races, 	
+	\item We have evaluated \sys on Linux 6.2, and find xxx data races, 	
 	with a false positive rate of xxx. Among these data races, xxx are 
@@
 Section~\ref{sec_technique} introduces our key techniques to address these 
-challenges. Section~\ref{sec_framework} introduces RaceMiner. 
+challenges. Section~\ref{sec_framework} introduces \sys. 
 Section~\ref{sec_evaluation} shows our evaluation. Section~\ref{sec_discussion} 
-makes a discussion about RaceMiner. Section~\ref{sec_related} presents related 
+makes a discussion about \sys. Section~\ref{sec_related} presents related 
 work, and Section~\ref{sec_conclusion} concludes this paper.

### 2023-07-19 2.motivation.tex
Add motivation

@@
 \label{sec_motivation}
+
+\subsection{A Motivating Example}
+\label{subsec_motivating_example}
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_bug_demo.pdf}
+	\figcaption{A null-pointer dereference due to data race in Linux 6.2.}
+	\label{fig_bug_demo}
+\end{figure}
+Figure~\ref{fig_bug_demo} shows a real null-pointer dereference caused by a 
+data race in the Linux DRM driver. In the DRM driver, the functions {\tt 
+exynos\_drm\_crtc\_atomic\_disable()} and {\tt exynos\_crtc\_handle\_event()} 
+can execute concurrently. We exploit \textcircled{\footnotesize{n}} to 
+represent the execution order of instructions and show one execution case in 
+the left of Figure~\ref{fig_bug_demo}. In Thread 1, the data structure fields 
+{\tt crtc->state->event} and {\tt crtc->state->active} is checked by an if 
+statement in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. After the 
+condition is calculated to be true, the function {\tt 
+exynos\_crtc\_handle\_event()} is executed in Thread 2. In this function, the 
+value of {\tt crtc->state->active} is assigned to {\tt event} 
+(\textcircled{\footnotesize{2}}) and then {\tt event} is checked in an if 
+statement (\textcircled{\footnotesize{3}}). If it is not NULL, the data 
+structure field {\tt crtc->state->event} is assigned with NULL 
+(\textcircled{\footnotesize{4}}). Right after this assignment, the function 
+{\tt drm\_crtc\_send\_vblank\_event} is called  in Thread 1 
+(\textcircled{\footnotesize{6}}) with the argument {\tt crtc->state->active}, 
+after acquiring the lock {\tt crtc->dev->event\_lock} 
+(\textcircled{\footnotesize{5}}). In the called function, the data structure 
+field {\tt crtc->state->event} is dereferenced through {\tt e->pipe} 
+(\textcircled{\footnotesize{6}}). In this execution case, the data structure 
+{\tt crtc->state->event} is first assigned with NULL and then dereferenced, and 
+thus a null-pointer dereference can occur.
+
+This bug is triggered only when {\tt crtc->state->event} is set to NULL by 
+Thread 2 right after the first condition of the if statement in Thread 1 is 
+calculated to be true. Such requirement is difficult to satisfy by executing 
+existing test suites. In fact, this bug had existed for nearly 6 years since 
+Linux 4.14 (Released in Nov. 2017), and it was fixed by us based on a report 
+generated by \sys.
+
+\subsection{Challenges}
+\label{subsec_challenges}
+Detecting data races and estimating their harmfulness in OS kernels have three 
+main challenges:
+
+\PP{C1: Getting locking rules.} The relationship between variables and locks 
+are not well documented in OS kernels, making it hard to determine whether a 
+specific data structure field should be protected by a lock and which lock is 
+required (locking rules), even for an expert developer. And thus existing 
+annotation-based approaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
+Anderson:PLDI09, Zhou:MICRO19, Flanagan:PASTE01, Flanagan:PLDI00, 
+Sadowski:PLATEAU14, ClangThreadSafety, Blackshear:OOPSLA18} are difficult to 
+apply to race detection in OS kernels. Other approaches~\cite{Choi:PLDI02, 
+Engler:SOSP03, Voung:FSE07, Pratikakis:PLDI06, Naik:PLDI06} employ 
+lockset-based analysis to detect data races automatically, but they do not 
+consider alias relations~\cite{Voung:FSE07, Engler:SOSP03} or just use 
+imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, 	
+Pratikakis:PLDI06, Naik:PLDI06}. However, due to the heavy use of pointers and 
+data structure fields in OS code, the alias relationships between variables can 
+be very complex, and thus lacking effective alias analysis can introduce many 
+false locking rules.
+
+\PP{C2: Dropping false data races.} Static analysis suffers from false 
+positives. For example, each data race involves more than one code paths that 
+should be able to concurrently execute. However, which code can execute 
+concurrently is not well documented for an OS kernel, and it is also hard to 
+determine concurrent code statically due to the complexity of OS code. Thus 
+static analysis can report many false data races.
+
+\PP{C3: Estimating harmfulness of data races.} Many data races are benign, and 
+can not cause memory or logic bugs, and thus developers are unwilling to put 
+effort into repairing them. To automatically detect harmful data races, most 
+approaches~\cite{Narayanasamy:PLDI07, Sen:PLDI08, Kasikci:SOSP13, 
+Kasikci:ASPLOS12} estimate harmfulness of data races through dynamic analysis, 
+and explore thread interleavings to trigger data races to estimate their 
+harmfulness. However, they suffer from low code coverage and thus can miss many 
+real harmful data races.

### 2023-07-21 1.introduction.tex
Update technique

@@
 benefits from precise field-sensitive alias relationships of alias graph, 
-\sys can accurately extract the accessed field and its corresponding 
+\sys can effectively extract the accessed field and its corresponding 
 protecting lock. After obtaining the accessed field and protecting lock pairs, 
@@
 (T2) \sys uses a {\em lock-usage analysis} to filter out false positives 
-caused by concurrency. After mining locking rules, \sys detects data 
-races by checking whether a given data structure field access violates the 
+validating concurrency of code paths. After mining locking rules, \sys detects 
+data races by checking whether a given data structure field access violates the 
 locking rules. To reduce false negatives, \sys conservatively assumes 

### 2023-07-21 3.technique.tex
Update technique

@@
 \label{sec_technique}
+To address the above challenges, we propose three key techniques. For {\em C1}, 
+we propose an {\em alias-aware rule mining method} to automatically deduce 
+locking rules. For {\em C2}, we propose a {\em lock-usage analysis} to filter 
+out false data races by validating concurrency of code paths. For {\em C3}, we 
+propose a pattern-based estimation to extract harmful races that can trigger 
+memory or logical bugs such as null-pointer dereference and data inconsistency. 
+We introduce them as follows:
+
+\subsection{Alias-Aware Rule Mining Method}
+\label{subsec_rule_mining}
+The relationship between variables and locks are not well documented in OS 
+kernels, but it can be inferred from the kernel code. Specifically, a data 
+structure field is often protected by the lock stored in the same data 
+structure. And thus if a data structure field is accessed after acquiring a 
+lock existing in the same data structure in most cases, the variable is likely 
+to be protected by the lock. Whether a data structure field and the protecting 
+lock exist in the same data structure can be determined though an alias 
+graph~\cite{Li:ASPLOS22, Kastrinis:CC18} by finding their common ancestor. 
+Based on this insight, we propose an {\em alias-aware rule mining method} to 
+deduce locking rules automatically. Moreover, with benefits from precise 
+field-sensitive alias relationships of alias graph, our alias-aware rule mining 
+method can find data structure filed and its protecting lock effectively.
+
+\PP{Alias Graph.} It is an important data structure to infer relationships 
+between data structure field and its protecting lock in our analysis, so we 
+introduce it and its update first.
+
+An alias graph is a 2-tuple $\mathit{G = \left<N, E\right>}$, where 
+$\mathit{N}$ is a set of nodes, and each node $\mathit{n}$ represents an alias 
+set that points to one abstract object. $\mathit{E}$ is a set of labeled edges. 
+Each edge is labeled with a data structure field or a dereference operator 
+``$\mathit{*}$'', which represents how an abstract object is accessed.
+
+An alias graph is updated by handling four types of instructions that  
+change alias relationships: MOVE($\mathit{v_1 = v_2}$), STORE ($\mathit{*v_2 = 
+v_1}$), LOAD ($\mathit{v_1 = *v_2}$) and GEP ({$\mathit{v_1 = \&v_1->f}$}). We 
+exploit {$\mathit{n_x}$} to represent the node whose representing alias set 
+includes $\mathit{v_x}$, and introduce how the four types of instructions 
+update alias graphs. For a MOVE operation, $\mathit{v_1}$ is moved from 
+$\mathit{n_1}$ to $\mathit{n_2}$. After this operation, $\mathit{v_1}$ and  
+$\mathit{v_2}$ are represented by the same node, which indicates they become 
+aliases. For a STORE operation, the existing outgoing edge from $\mathit{n_2}$ 
+is dropped first, and then a new edge labeled with $\mathit{*}$ from 
+$\mathit{n_2}$ to $\mathit{n_1}$ is inserted. After this operation, 
+$\mathit{v_1}$ and $\mathit{*v_2}$ are represented by the same node, which 
+indicates they become aliases. For a LOAD operation, the analysis first finds 
+the destination node of the edge that comes from $\mathit{n_2}$ and is labeled 
+with $\mathit{*}$, and then move $\mathit{v_1}$ to the destination node. And 
+after this operation, $\mathit{v_1}$ and $\mathit{*v_2}$ are represented by the 
+same node, which indicates they become aliases. GEP operation is similar to 
+LOAD, expect that the edge is labeled with a data structure field $\mathit{f}$, 
+instead of a dereference operator $\mathit{*}$.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{figures/fig_alias_graph_demo.pdf}
+	\figcaption{Example of alias graph.}
+	\label{fig_alias_graph_demo}
+\end{figure}
+
+\noindent{\textbf{\em Example.}} Figure~\ref{fig_alias_graph_demo} shows a 
+piece of driver-like source code and its alias graph. In this example, after an 
+GEP (\&dev->lock) operation at Line 6, an edge labeled with {\tt lock} from 
+node $\mathit{n_1}$ to node $\mathit{n_4}$ is inserted. Similarly, an edge 
+labeled with {\tt flag} from node $\mathit{n_1}$ to node $\mathit{n_2}$ is 
+inserted after Line 7. At last, an edge labeled with a dereference operator 
+($\mathit{*}$) is inserted after the STORE (*fp = f) operation at Line 8. In 
+this example, {\tt \&dev->flag} is represented by node $\mathit{n_2}$, and {\tt 
+\&dev->lock} is represented by node $\mathit{n_4}$. The two nodes have a common 
+ancestor $\mathit{n_1}$, and thus the accessed data structure field {\tt 
+\&dev->flag} and the protecting lock {\tt \&dev->lock} can be inferred to exist 
+in the same data structure (namely {\tt Dev}).

### 2023-07-23 1.introduction.tex
update

@@
 Flanagan:PLDI00, Sadowski:PLATEAU14, ClangThreadSafety, Blackshear:OOPSLA18} 
-rely on developers to 
-supply annotations that describe the locking discipline such as which lock 
-protects a specific data structure field, and detect data races by exploiting 
-data structure field accesses that violate these disciplines. For example, 
-Clang thread safety analysis~\cite{ClangThreadSafety} requires developers to 
-label a variable and the lock to protect it with a {\tt GUARDED\_BY} attribute. 
-And then detects data races by finding variable accesses that violate the given 
+rely on developers to supply annotations that describe the locking discipline 
+such as which lock protects a specific variable, and detect data races by 
+exploiting variable accesses that violate these disciplines. For example, Clang 
+thread safety analysis~\cite{ClangThreadSafety} requires developers to label a 
+variable and the lock to protect it with a {\tt GUARDED\_BY} attribute. And 
+then detects data races by finding variable accesses that violate the given 
 attribute. However, such annotation-based approaches are not suitable for race 
@@
 and then use the mined rules to detect related bugs. For example, 
-LockDoc~\cite{Lochmann:EuroSys19} records accesses to data structure fields and 
-lock acquisitions of an instrumented Linux kernel, and then infers locking 
-rules from the recorded accesses. After that, LockDoc automatically locates 
-data structure fields access that violate the inferred locking rules to detect 
-concurrency bugs such as data races. However, dynamic analysis suffers from low 
-code coverage, and thus the locking rules inferred through execution traces can 
-be imprecise. Besides, dynamic analyses are hard to deploy, because they need 
-to run OS kernels with instruments. At last, it is difficult for dynamic 
-tools to estimate the harmfulness of a data race, because race conditions are 
-hard to trigger by running the checked software~\cite{Fonseca:DSN10, 
+LockDoc~\cite{Lochmann:EuroSys19} records accesses to variables and lock 
+acquisitions of an instrumented Linux kernel, and then infers locking rules 
+from the recorded accesses. After that, LockDoc automatically locates variables 
+access that violate the inferred locking rules to detect concurrency bugs such 
+as data races. However, dynamic analysis suffers from low code coverage, and 
+thus the locking rules inferred through execution traces can be imprecise. 
+Besides, dynamic analyses are hard to deploy, because they need to run OS 
+kernels with instruments. At last, it is difficult for dynamic tools to 
+estimate the harmfulness of a data race, because race conditions are hard to 
+trigger by running the checked software~\cite{Fonseca:DSN10, 
 Burckhardt:ASPLOS10, Liu:FSE14, Zhou:EASE15}.
@@
 (T1) \sys uses an {\em alias-aware rule mining method} to automatically 
-deduce locking rules. We observe that an access to a data structure field is 
-often protected by the lock in the same data structure as the accessed field. 
-And this relationship between accessed field and the protecting lock can be 
+deduce locking rules. We observe that an access to a variable is often 
+protected by the lock in the same data structure as the accessed field. And 
+this relationship between accessed field and the protecting lock can be 
 effectively inferred from a special data structure named alias 
@@
 validating concurrency of code paths. After mining locking rules, \sys detects 
-data races by checking whether a given data structure field access violates the 
+data races by checking whether a given variable access violates the 
 locking rules. To reduce false negatives, \sys conservatively assumes 

### 2023-07-23 2.motivation.tex
Update technique
update

@@
 % !TeX spellcheck = en_US
-\section{Motivation}
+\section{Background and Motivation}
 \label{sec_motivation}
@@
 represent the execution order of instructions and show one execution case in 
-the left of Figure~\ref{fig_bug_demo}. In Thread 1, the data structure fields 
-{\tt crtc->state->event} and {\tt crtc->state->active} is checked by an if 
-statement in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. After the 
-condition is calculated to be true, the function {\tt 
-exynos\_crtc\_handle\_event()} is executed in Thread 2. In this function, the 
-value of {\tt crtc->state->active} is assigned to {\tt event} 
-(\textcircled{\footnotesize{2}}) and then {\tt event} is checked in an if 
-statement (\textcircled{\footnotesize{3}}). If it is not NULL, the data 
-structure field {\tt crtc->state->event} is assigned with NULL 
+the left of Figure~\ref{fig_bug_demo}. In Thread 1, the variables {\tt 
+crtc->state->event} and {\tt crtc->state->active} is checked by an if statement 
+in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. After the condition 
+is calculated to be true, the function {\tt exynos\_crtc\_handle\_event()} is 
+executed in Thread 2. In this function, the value of {\tt crtc->state->active} 
+is assigned to {\tt event} (\textcircled{\footnotesize{2}}) and then {\tt 
+event} is checked in an if statement (\textcircled{\footnotesize{3}}). If it is 
+not NULL, the variable {\tt crtc->state->event} is assigned with NULL 
 (\textcircled{\footnotesize{4}}). Right after this assignment, the function 
@@
 after acquiring the lock {\tt crtc->dev->event\_lock} 
-(\textcircled{\footnotesize{5}}). In the called function, the data structure 
-field {\tt crtc->state->event} is dereferenced through {\tt e->pipe} 
+(\textcircled{\footnotesize{5}}). In the called function, the variable {\tt 
+crtc->state->event} is dereferenced through {\tt e->pipe} 
 (\textcircled{\footnotesize{6}}). In this execution case, the data structure 
@@
 are not well documented in OS kernels, making it hard to determine whether a 
-specific data structure field should be protected by a lock and which lock is 
+specific variable should be protected by a lock and which lock is 
 required (locking rules), even for an expert developer. And thus existing 

### 2023-07-23 3.technique.tex
update

@@
 The relationship between variables and locks are not well documented in OS 
-kernels, but it can be inferred from the kernel code. Specifically, a data 
-structure field is often protected by the lock stored in the same data 
-structure. And thus if a data structure field is accessed after acquiring a 
-lock existing in the same data structure in most cases, the variable is likely 
-to be protected by the lock. Whether a data structure field and the protecting 
-lock exist in the same data structure can be determined though an alias 
-graph~\cite{Li:ASPLOS22, Kastrinis:CC18} by finding their common ancestor. 
-Based on this insight, we propose an {\em alias-aware rule mining method} to 
-deduce locking rules automatically. Moreover, with benefits from precise 
-field-sensitive alias relationships of alias graph, our alias-aware rule mining 
-method can find data structure filed and its protecting lock effectively.
+kernels, but it can be inferred from the kernel code. Specifically, a variable 
+is often protected by the lock stored in the same data structure. And thus if a 
+variable is accessed after acquiring a lock existing in the same data structure 
+in most cases, it is likely to be protected by the lock. Whether a variable and 
+the protecting lock exist in the same data structure can be determined though 
+an alias graph~\cite{Li:ASPLOS22, Kastrinis:CC18} by finding their common 
+ancestor. Based on this insight, we propose an {\em alias-aware rule mining 
+method} to deduce locking rules automatically. Moreover, with benefits from 
+precise field-sensitive alias relationships of alias graph, our alias-aware 
+rule mining method can find data structure filed and its protecting lock 
+effectively.
 
 \PP{Alias Graph.} It is an important data structure to infer relationships 
-between data structure field and its protecting lock in our analysis, so we 
-introduce it and its update first.
+between variable and its protecting lock in our analysis, so we introduce it 
+and its update first.
 
@@
 Each edge is labeled with a data structure field or a dereference operator 
-``$\mathit{*}$'', which represents how an abstract object is accessed.
+``$\mathit{*}$'', which represents how an abstract object is accessed. In an 
+alias node, a variable residing in a node followed by a sequence of edge
+labels form an access path~\cite{Kastrinis:CC18, Cheng:PLDI00}. In this paper, 
+we replace the variable in an access path with its structure name to represent 
+a data structure field.
 
@@
 inserted after Line 7. At last, an edge labeled with a dereference operator 
-($\mathit{*}$) is inserted after the STORE (*fp = f) operation at Line 8. In 
-this example, {\tt \&dev->flag} is represented by node $\mathit{n_2}$, and {\tt 
-\&dev->lock} is represented by node $\mathit{n_4}$. The two nodes have a common 
-ancestor $\mathit{n_1}$, and thus the accessed data structure field {\tt 
-\&dev->flag} and the protecting lock {\tt \&dev->lock} can be inferred to exist 
-in the same data structure (namely {\tt Dev}).
+($\mathit{*}$) is inserted after the STORE (*fp = f) operation at Line 8. The 
+final alias graph is shown in Figure~\ref{fig_alias_graph_demo}(b), and access 
+paths are shown in the bottom left corner. Take node $\mathit{n_3}$ as an 
+example, it can represent two fields, one is {\tt Dev.flag*}, and the other is 
+{\tt int\_pointer*} (we exploit int\_pointer to represent a pointer points to 
+an integer, and regard it as a data structure for convenience).
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_pseudocode_get_access.pdf}
+	\figcaption{Pseudocodes to get accessed field and protecting lock.}
+	\label{fig_pseudocode_get_access}
+\end{figure}
+
+Given an alias graph, whether a variable and a lock exist in the same data 
+structure can be determined by finding a common ancestor. If they are in the 
+same data structure, the lock is likely to protect the variable when it is 
+accessed. Figure~\ref{fig_pseudocode_get_access} shows the pseudocode to 
+get the field of the accessed variable and the field of the protecting lock in 
+the form of access path, if they exist in the same data structure. Given a node 
+of an accessed variable and a node of a lock variable, the analysis first gets 
+the common ancestor of the two nodes (Line 1). And then, if the common ancestor 
+does not exist, the analysis returns NULL (Lines 2-3). Otherwise, the analysis 
+gets the access path for the node of the accessed variable and the node of the 
+lock variable from the common ancestor (Lines 5-7).
+
+%\begin{table}[htbp]
+%	\tablecaption{Notation table of pseudocodes.}
+%	\label{tbl_notation}
+%	\centering
+%	\scriptsize
+%	\begin{spacing}{0.98}
+%		\begin{tabular}{@{ }p{2.7cm}p{5.05cm}@{ }}
+%		\toprule[1pt]
+%		GetAliasNode($\mathit{var}$, $\mathit{alias\_graph}$) & Get the node 
+%		the variable $\mathit{var}$ exists in from the alias graph 
+%		$\mathit{alias\_graph}$. \\ \hline
+%		GetCommonAncestor($\mathit{node_1}$, $\mathit{node_2}$) & Get the 
+%		common ancestor of the two nodes $\mathit{node_1}$ and 
+%		$\mathit{node_2}$. \\ \hline
+%		GetAccessPath($\mathit{node_1}$, $\mathit{node_2}$) & Get the access 
+%		path from $\mathit{node_1}$ to $\mathit{node_2}$ \\ \hline
+%		\bottomrule[1pt]	
+%		\end{tabular}
+%	\end{spacing}
+%\end{table}
+
+Take the alias graph in Figure~\ref{fig_alias_graph_demo} as an example, {\tt 
+\&dev->flag} is represented by node $\mathit{n_2}$, and {\tt \&dev->lock} is 
+represented by node $\mathit{n_4}$. The two nodes have a common ancestor 
+$\mathit{n_1}$, and thus the accessed variable {\tt \&dev->flag} and the 
+protecting lock {\tt \&dev->lock} can be inferred to exist in the same data 
+structure (namely {\tt Dev}). Therefore, the structure field Dev.flag is likely 
+to be protected the lock stored in the structure field Dev.lock.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_workflow.pdf}
+	\figcaption{Workflow of locking-rule mining.}
+	\label{fig_workflow}
+\end{figure}
+
+Based on the alias graph, our alias-aware rule mining method performs an 
+inter-procedural path-based~\cite{Li:ASPLOS22}, field-sensitive and alias-aware 
+analysis to effectively mine locking rules about whether a specific variable 
+should be protected by a lock and which lock is required. Overall, our 
+alias-aware rule mining method has three main stages shown in 
+Figure~\ref{fig_workflow}. In Stage 1, it extract key variable that may be 
+protected by a lock. In Stage 2, it collect all accesses to key fields as well 
+as code paths these accesses exist in. In Stage 3, it deduce locking rules by 
+calculate the proportion of field accesses protected by a specific lock in all 
+field accesses.
+
+\PP{S1: Key-field Extraction.} The OS Kernel has a large code base with 
+numerous variables. However, only a small part of variables should be protected 
+by a specific lock, and thus collecting all variable accesses can introduce 
+much unnecessary overhead. Generally, only a few field accesses can miss 
+necessary protecting lock, due to carelessness of developers. Based on this 
+insight, our analysis first extracts key fields that may need to be protected 
+by a specific lock, by performing a lock-set analysis to find whether a given 
+variable is accessed after acquiring a lock in the same data structure as the 
+accessed field in any code path.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{figures/fig_pseudocode_lock_set.pdf}
+	\figcaption{Pseudocodes of lock-set analysis.}
+	\label{fig_pseudocode_lock_set}
+\end{figure}
+
+Figure~\ref{fig_pseudocode_lock_set} shows the lock-set analysis based on alias 
+graph. The analysis first gets the operand of the instruction (Line 1), and 
+then get the node of the operand from the alias graph (Line 2). If the 
+instruction is a lock operation, the node of the operand is inserted into the 
+lock set (Lines 3-4). Otherwise, if the instruction is an unlock operation, the 
+node of the operand is removed from the lock set (Lines 5-6).
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_pseudocode_field_extract.pdf}
+	\figcaption{Pseudocodes of key-field extraction.}
+	\label{fig_pseudocode_field_extract}
+\end{figure}
+
+Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to collect key 
+fields that may be protected by a specific lock when are accessed, based on 
+lock-set analysis and alias graph. The analysis start analyzing
+
+For each function without a caller function, the analysis

### 2023-07-23 paper.tex
Add motivation
Update technique
update

@@
 \usepackage{xspace}
+\usepackage{xstring}
 
@@
 
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+
 \newcommand{\sys}{\mbox{RaceMiner}\xspace}
+
+\newcommand{\figcaption}{% 
+	\setlength{\abovecaptionskip}{4pt}%	\setlength{\belowcaptionskip}{1pt}% 
+	\caption}
+
+\newcommand{\tablecaption}{% 
+	\setlength{\abovecaptionskip}{6pt}%
+	\setlength{\belowcaptionskip}{2pt}% 
+	\caption}
+
+\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
+
+\newcommand{\PP}[1]{
+	\vspace{2px}
+	\noindent{\bf \IfEndWith{#1}{.}{#1}{#1.}}
+}
 
@@
 data, fine-grained locking mechanisms are introduced. However, it is hard to 
-determine whether a lock should be used for a specific data structure field 
-(locking rules) even for an expert developer, due to poor documentation and 
-complicated logic of OS code. As a result, OS kernel is prone to data races, 
-because necessary locks may be missed by mistake. Static analysis is a common 
-technique to help improve code quality, but it is quite challenging to detect 
-data races in OS kernels automatically, because of lack of knowledge of locking 
-rules and high complexity of concurrent execution.
+determine whether a lock should be used for a specific variable (locking rules) 
+even for an expert developer, due to poor documentation and complicated logic 
+of OS code. As a result, OS kernel is prone to data races, because necessary 
+locks may be missed by mistake. Static analysis is a common technique to help 
+improve code quality, but it is quite challenging to detect data races in OS 
+kernels automatically, because of lack of knowledge of locking rules and high 
+complexity of concurrent execution.
 

### 2023-07-24 3.technique.tex
update

@@
 the common ancestor of the two nodes (Line 1). And then, if the common ancestor 
-does not exist, the analysis returns NULL (Lines 2-3). Otherwise, the analysis 
-gets the access path for the node of the accessed variable and the node of the 
-lock variable from the common ancestor (Lines 5-7).
+does not exist, the analysis returns a NULL pair (Lines 2-3). Otherwise, the 
+analysis gets the access path for the node of the accessed variable and the 
+node of the lock variable from the common ancestor (Lines 5-7).
 
@@
 
-\PP{S1: Key-field Extraction.} The OS Kernel has a large code base with 
+\PP{S1: Key-field extraction.} The OS Kernel has a large code base with 
 numerous variables. However, only a small part of variables should be protected 
@@
 Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to collect key 
-fields that may be protected by a specific lock when are accessed, based on 
-lock-set analysis and alias graph. The analysis start analyzing
+fields that may be protected by a specific lock, based on lock-set analysis and 
+the alias graph. The analysis starts from each function without a caller 
+function (Lines 36-40) and performs a path-based analysis~\cite{Li:ASPLOS22} 
+(Lines 26-31). For each instruction in the code path, the analysis first update 
+the alias graph according to the instruction with the four operations (MOVE, 
+STORE, LOAD and GEP) (Line 1), and then performs a lock-set analysis (Line 2) 
+to get all locks acquired. After update the alias graph and the lock set, the 
+analysis gets the node of the operand from the new alias graph (Lines 3-4), and 
+then update the set of key fields, according to the type of instructions (Lines 
+5-24). Specifically, if the instruction is a write or a read, for each node of 
+lock variable in the lock set, the analysis uses the alias graph to extract the 
+protected data structure field (Lines 9-11). If the field (in the form of 
+access path) is not NULL, it is inserted into the set of key fields. As for a 
+function call instruction, the analysis first analyzes the called function with 
+the alias graph and lock set of the caller function (Lines 18-21), and then 
+merges the returned key fields to the caller function (Line 22).
 
-For each function without a caller function, the analysis
+\PP{S2: Field-access collection.}

### 2023-07-25 3.technique.tex
update technique

--- 2023-07-23 3.technique.tex
@@
 the common ancestor of the two nodes (Line 1). And then, if the common ancestor 
-does not exist, the analysis returns NULL (Lines 2-3). Otherwise, the analysis 
-gets the access path for the node of the accessed variable and the node of the 
-lock variable from the common ancestor (Lines 5-7).
-
-%\begin{table}[htbp]
-%	\tablecaption{Notation table of pseudocodes.}
-%	\label{tbl_notation}
-%	\centering
-%	\scriptsize
-%	\begin{spacing}{0.98}
-%		\begin{tabular}{@{ }p{2.7cm}p{5.05cm}@{ }}
-%		\toprule[1pt]
-%		GetAliasNode($\mathit{var}$, $\mathit{alias\_graph}$) & Get the node 
-%		the variable $\mathit{var}$ exists in from the alias graph 
-%		$\mathit{alias\_graph}$. \\ \hline
-%		GetCommonAncestor($\mathit{node_1}$, $\mathit{node_2}$) & Get the 
-%		common ancestor of the two nodes $\mathit{node_1}$ and 
-%		$\mathit{node_2}$. \\ \hline
-%		GetAccessPath($\mathit{node_1}$, $\mathit{node_2}$) & Get the access 
-%		path from $\mathit{node_1}$ to $\mathit{node_2}$ \\ \hline
-%		\bottomrule[1pt]	
-%		\end{tabular}
-%	\end{spacing}
-%\end{table}
+does not exist, the analysis returns a NULL pair (Lines 2-3). Otherwise, the 
+analysis gets the access path for the node of the accessed variable and the 
+node of the lock variable from the common ancestor (Lines 5-7).
 
@@
 
-Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to collect key 
-fields that may be protected by a specific lock when are accessed, based on 
-lock-set analysis and alias graph. The analysis start analyzing
+Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to extract key 
+fields that may be protected by a specific lock, based on lock-set analysis and 
+the alias graph. The analysis starts from each function without a caller 
+function (Lines 16-26) and performs a path-based analysis~\cite{Li:ASPLOS22} 
+(Lines 17-25). For each instruction in the code path, the analysis first update 
+the alias graph according to the instruction with the four operations (MOVE, 
+STORE, LOAD and GEP) (Line 1), and then performs a lock-set analysis (Line 2) 
+to get all acquired locks. After updating the alias graph and the lock set, if 
+the instruction is a write or a read, the analysis first gets the node of the 
+operand with the new alias graph (Lines 4-5). And then, for each node of 
+lock variable in the lock set, the analysis uses the alias graph to extract the 
+protected data structure field (Lines 6-12). If the field is not NULL, it is 
+inserted into the set of key fields (Lines 9-11). For inter-procedural 
+analysis, the analysis copies function definition into its call sites and this 
+operation is not shown in Figure~\ref{fig_pseudocode_field_extract} for 
+convenience.
 
-For each function without a caller function, the analysis
+\PP{S2: Field-access collection.} With the key fields extracted in Stage 1, the 
+analysis can only collects accesses to the variables that is protected in some 
+cases, because if a variable is never protected by any lock when is accessed, 
+it is less likely to be shared by different threads, and thus can not introduce 
+no concurrency issue.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_pseudocode_access_collect.pdf}
+	\figcaption{Pseudocodes of key-field extraction.}
+	\label{fig_pseudocode_access_collect}
+\end{figure}
+
+Figure~\ref{fig_pseudocode_access_collect} shows the pseudocode to collect all 
+accesses to key fields. Similarly to the key-field extraction, this stage also 
+performs a path-based analysis. For each instruction in the code path (Lines 
+39-45), the analysis first updates the alias graph and the lock set according 
+to the handled instruction (Lines 1-2). And then, the access type (either a 
+write or a read) is set according to the instruction (Lines 5-8). However, if 
+the instruction is not an access operation, the function returns with a NULL 
+value for the field access (Line 10). Otherwise, the analysis first gets the 
+node of the operand with the new alias graph (Lines 12-13), and then for each 
+node of the lock variable in the lock set, the analysis uses the alias graph 
+toextract the fields that the accessed variable and the lock variable stored in 
+(Lines 16-17). If the fields are found, the field access is returned with the 
+found fields (Lines 18-22). If the fields are not found, the accessed variable  
+does not protected by any lock. In this case, if the accessed variable is 
+stored in a key field, the field access is returned with a NULL lock variable 
+(Lines 25-30).
+
+\PP{S3: Locking-rule deduction.} After collecting all accesses to key fields, 
+the analysis deduces locking rules based on statistics. However, on the one 
+hand, distinguishing different accesses to the same data structure field by 
+different program sites is not fine-grained enough, because the access to a 
+variable and the acquirement to a specific lock are often packaged in a special 
+function, and all accesses under different calling context are regarded as the 
+same access. On the other hand, distinguishing different accesses by different 
+code paths can also suffer from inaccuracy when a function contains many branch 
+statement and cause numerous code paths for the same variable access. Based on 
+these consideration, the analysis distinguishes different accesses to the same 
+data structure field by different calling contexts. Specifically, given a key 
+field {\em f} and a lock {\em l}, the analysis first finds all access to {\em 
+f} with lock {\em l} from the collected field accesses in Stage 2, and gets the 
+number {\em locked\_access\_num} by counting different calling contexts 
+extracted from the code paths of these field accesses. Then, the analysis finds 
+all access to {\em f} (no matter which lock is accessed), and gets the number 
+{\em all\_access\_num} in the same way as {\em locked\_access\_num}. If the 
+ratio {\em locked\_access\_num}/{\em all\_access\_num} is larger than a given 
+threshold, and these is at least one write access to {\em f}, the data 
+structure field {\em f} is inferred to be protected by the lock {\em l}.

### 2023-07-25 2.motivation.tex
update technique

@@
 	\centering
-	\includegraphics[width=1\linewidth]{figures/fig_bug_demo.pdf}
+	\includegraphics[width=1\linewidth]{figures/fig_demo_bug.pdf}
 	\figcaption{A null-pointer dereference due to data race in Linux 6.2.}
-	\label{fig_bug_demo}
+	\label{fig_demo_bug}
 \end{figure}
-Figure~\ref{fig_bug_demo} shows a real null-pointer dereference caused by a 
+Figure~\ref{fig_demo_bug} shows a real null-pointer dereference caused by a 
 data race in the Linux DRM driver. In the DRM driver, the functions {\tt 
@@
 represent the execution order of instructions and show one execution case in 
-the left of Figure~\ref{fig_bug_demo}. In Thread 1, the variables {\tt 
+the left of Figure~\ref{fig_demo_bug}. In Thread 1, the variables {\tt 
 crtc->state->event} and {\tt crtc->state->active} is checked by an if statement 

### 2023-07-25 3.technique.tex
update technique

@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{figures/fig_alias_graph_demo.pdf}
+	\includegraphics[width=0.9\linewidth]{figures/fig_demo_alias_graph.pdf}
 	\figcaption{Example of alias graph.}
-	\label{fig_alias_graph_demo}
+	\label{fig_demo_alias_graph}
 \end{figure}
 
-\noindent{\textbf{\em Example.}} Figure~\ref{fig_alias_graph_demo} shows a 
+\noindent{\textbf{\em Example.}} Figure~\ref{fig_demo_alias_graph} shows a 
 piece of driver-like source code and its alias graph. In this example, after an 
@@
 ($\mathit{*}$) is inserted after the STORE (*fp = f) operation at Line 8. The 
-final alias graph is shown in Figure~\ref{fig_alias_graph_demo}(b), and access 
+final alias graph is shown in Figure~\ref{fig_demo_alias_graph}(b), and access 
 paths are shown in the bottom left corner. Take node $\mathit{n_3}$ as an 
@@
 
-Take the alias graph in Figure~\ref{fig_alias_graph_demo} as an example, {\tt 
+Take the alias graph in Figure~\ref{fig_demo_alias_graph} as an example, {\tt 
 \&dev->flag} is represented by node $\mathit{n_2}$, and {\tt \&dev->lock} is 
@@
 structure field {\em f} is inferred to be protected by the lock {\em l}.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_demo_rule_mining.pdf}
+	\figcaption{Example of locking-rule mining.}
+	\label{fig_demo_rule_mining}
+\end{figure}
+
+\noindent{\textbf{\em Example.}} We use an example in 
+Figure~\ref{fig_demo_rule_mining} to illustrate how to mine locking rules 
+with the three stages. The analysis first perform a path-based analysis to 
+extract key fields. Take the code path {\tt Line 12, Line 6, Line 7, Line 8, 
+Line 9} as an example. The analysis updates the alias graph by handling two GEP 
+operations ({\em \&dev->lock} and {\em fp = dev->flag}) and a STORE operation 
+({\em *fp = f}). The final alias graph is shown in 
+Figure~\ref{fig_demo_rule_mining}(b). At the same time of alias analysis, the 
+analysis records the acquired lock {\em Dev.lock} into the lock set after the 
+lock instruction at Line 6. When analyzing the read instruction at Line 7, the 
+analysis finds the node of the accessed field ($\mathit{n_2}$) and the node of 
+the lock stored in the lock set ($\mathit{n_4}$) has the same ancestor 
+($\mathit{n_1}$), and thus {\em Dev.flag} is a key field. Similarly, {\em 
+Dev.flag*} is also a key field, due to the write instruction at Line 8.
+
+After extracting key fields, the method collects all accesses to key fields 
+with lock set analysis and the alias graph. Take the code path {\tt Line 15, 
+Line 6, Line 7, Line 8, Line 9, Line 16 and Line 17} as an example, through a 
+lock instruction at Line 6, the analysis records the acquired lock {\em 
+Dev.lock} into the lock set. When analyzing the read instruction at Line 7, the 
+analysis finds the node of the accessed field ($\mathit{n_2}$) and the node of 
+the lock stored in the lock set ($\mathit{n_4}$) has the same ancestor 
+($\mathit{n_1}$), and thus records a field access <{\em Dev.flag}, {\em 
+Dev.lock}>. Similar to the read instruction at Line 7, the method also records 
+a field access <{\em Dev.flag*}, {\em Dev.lock}> for the write instruction at 
+Line 8. Then through an unlock operation at Line 9, the analysis removes the 
+lock {\em Dev.lock} is removed from the lock set. As a result, the key fields 
+{\em Dev.flag} and {\em Dev.flag*} are accessed without acquiring any lock, and 
+thus the analysis records two field accesses <{\em Dev.flag}, {\em NULL}> and 
+<{\em Dev.flag*}, {\em NULL}>. The final field accesses are shown in 
+Figure~\ref{fig_demo_rule_mining}.
+
+After collecting all field accesses, the analysis finds that the fields {\em 
+Dev.flag} and {\em Dev.flag*} are accessed under three calling contexts, and 
+two of them are protected by the lock {\em Dev.lock}. And thus the two fields 
+{\em Dev.flag} and {\em Dev.flag*} is deduce to be protected by the lock {\em 
+Dev.lock} (assume the threshold of the ratio locked\_access\_num / 
+all\_access\_num is 0.6).
+
+\subsection{Lock-Usage Analysis}
+\label{subsec_lock_usage_analysis}

### 2023-07-26 3.technique.tex
Update technique

@@
 threshold, and these is at least one write access to {\em f}, the data 
-structure field {\em f} is inferred to be protected by the lock {\em l}.
+structure field {\em f} is inferred to be protected by the lock {\em l}. After 
+deducting locking rules, the analysis detecting data races by checking whether 
+a field access validates any locking rules.
 
@@
 Line 8. Then through an unlock operation at Line 9, the analysis removes the 
-lock {\em Dev.lock} is removed from the lock set. As a result, the key fields 
-{\em Dev.flag} and {\em Dev.flag*} are accessed without acquiring any lock, and 
-thus the analysis records two field accesses <{\em Dev.flag}, {\em NULL}> and 
-<{\em Dev.flag*}, {\em NULL}>. The final field accesses are shown in 
-Figure~\ref{fig_demo_rule_mining}.
+lock {\em Dev.lock} from the lock set. As a result, the key fields {\em 
+Dev.flag} and {\em Dev.flag*} are accessed without acquiring any lock, and thus 
+the analysis records two field accesses <{\em Dev.flag}, {\em NULL}> and <{\em 
+Dev.flag*}, {\em NULL}>. The final field accesses are shown in 
+Figure~\ref{fig_demo_rule_mining}(c).
 
@@
 {\em Dev.flag} and {\em Dev.flag*} is deduce to be protected by the lock {\em 
-Dev.lock} (assume the threshold of the ratio locked\_access\_num / 
-all\_access\_num is 0.6).
+Dev.lock} (assume the threshold of the ratio {\em locked\_access\_num} / 
+{\em all\_access\_num} is 0.6). However, the accesses at Lines 16 and 17 are 
+not protected by {\em Dev.lock} and thus causing two data races.
 
@@
 \label{subsec_lock_usage_analysis}
+Our analysis takes functions that has no caller function as the entry as 
+existing work~\cite{Li:ASPLOS22}, and performs a path-based analysis. It 
+assumes that all codes reaching from entry functions can execute concurrently 
+to reduce false negatives. However, this assumption is too conservative and can 
+introduce many false positives because not all entry functions can execute 
+concurrently in fact. We observe that each kernel module has an initialization 
+phase, after which many function can be called concurrently by other modules. 
+When performing initialization, the kernel module serially initialized the lock 
+and prepare other data for subsequent operations. And thus functions with the 
+lock initialization and functions called by them tend not to be executed 
+concurrently. Based on the observation, we propose a lock-usage analysis to 
+filter out false positives caused by code that can not execute concurrently.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{figures/fig_pseudocode_lock_usage.pdf}
+	\figcaption{Pseudocodes of lock-usage analysis.}
+	\label{fig_pseudocode_lock_usage}
+\end{figure}
+
+Our lock-usage analysis uses the union-find set~\cite{Galler:ACM64} to get all 
+functions that reachable from the lock initialization function such as {\em 
+mutex\_init()} and {\em spin\_lock\_init()}. 
+Figure~\ref{fig_pseudocode_lock_usage} shows the pseudocodes of the lock-usage 
+analysis. For each analyzed function in the OS kernel, the analysis first gets 
+all the called function of it (Lines 3-4), and then union the analyzed function 
+and the called function to update the union-find set (Line 5). After processing 
+all the call instructions, the analysis judging whether a function executes in 
+the initialization phase by checking whether it exist in the same set with the 
+given lock initialization function (Lines 9-11).
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{figures/fig_demo_lock_usage.pdf}
+	\figcaption{A false data race filtered out by our lock-usage analysis.}
+	\label{fig_demo_lock_usage}
+\end{figure}
+
+\noindent{\textbf{\em Example.}} Figure~\ref{fig_demo_lock_usage} shows a false 
+data race filtered out by our lock-usage analysis in the Linux DRM scheduler. 
+For example in {\em drm\_sched\_entity\_set\_priority()}, the accesses to {\em 
+entity->priority} is protected by the lock {\em entity->rq\_lock} in most 
+cases. Therefore, {\em entity-priority} is deduced to be protected by the lock 
+{\em entity->rq\_lock}. But in {\em drm\_sched\_entity\_init()}, {\em 
+entity->priority} is written without acquiring the lock {\em entity->rq\_lock} 
+and thus introducing a possible data race. However, the lock initialization 
+function {\em spin\_lock\_init()} is called at Line 86 by {\em 
+drm\_sched\_entity\_init()}. And thus {\em drm\_sched\_entity\_init()} is 
+inferred to execute in the module initialization phase and can not execute 
+concurrently with other function, and this is a false data race.
+
+\subsection{Pattern-Based Estimation}
+\label{subsec_estimation}

### 2023-07-27 3.technique.tex
Finish the draft of technique

@@
 \label{subsec_estimation}
+Many data races are benign or introduced by developers deliberately to improve 
+the kernel performance. They can not cause memory or logic bugs, and thus 
+developers are unwilling to put effort into repairing them. We observe that 
+harmful data races match some typical patterns, and propose a {\em 
+pattern-based estimation} to extract data races that can cause memory or logic 
+bugs.
+
+At present, we propose four patterns that can cause null-pointer dereference, 
+infinite loop, data inconsistency and unprotected write, because these four 
+patterns are common and dangerous in OSes. The four patterns are shown in 
+Figure~\ref{fig_pattern}, and we assume that accesses to the fields of {\em 
+dev} should be protected by {\em dev->lock}.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.85\linewidth]{figures/fig_pattern.pdf}
+	\figcaption{Three harmful patterns of data races.}
+	\label{fig_pattern}
+\end{figure}
+
+\begin{itemize}
+	\item \PP{Null-pointer dereference.} The two accesses to {\em dev->event} 
+	are not protected by {\em dev->lock}. This can cause a null-pointer 
+	dereference if {\em dev->event} is set to NULL right after the condition of 
+	the if statement is checked to be true. To recognize such pattern, the 
+	analysis first locates the data race, if it is checked by an if statement 
+	and then dereferenced, a possible null-pointer dereference can occur.
+	\item \PP{Infinite loop.} The condition of the while statement does not 
+	protected by {\em dev->lock}. This can cause a infinite loop when {\em 
+	dev->cnt} is assigned with a non-zero value every time before it is checked 
+	by the while statement. The analysis detects such bugs by judge whether the 
+	data race acts as a condition of a loop statement.
+	\item \PP{Data inconsistency.} Two different fields of the same data 
+	structure are accessed without acquiring {\em dev->lock}. This can cause a 
+	data inconsistency when the data structure field {\em dev->clock} is 
+	changed by another thread right after the access to {\em dev->clock->min}. 
+	To recognize such pattern, the analysis detect whether multiple fields of 
+	the same data structure are accessed without acquiring the protecting lock.
+	\item \PP{Unprotected write.} The write access to {\em dev->event} is not 
+	protected by {\em dev->lock}. This is dangerous because the value {\em 
+	dev->event} can be modified at any time when other threads access it, and 
+	thus introduce unpredictable behavior. The analysis detect such pattern by 
+	judge whether a data race occurs in a write access.
+\end{itemize}

### 2023-07-27 4.framework.tex
Finish the draft of technique
Finish the draft of framework

@@
 % !TeX spellcheck = en_US
-\section{Framework}
+\section{\sys Approach}
 \label{sec_framework}
+
+Based on the three key techniques in Section~\ref{sec_technique}, we develop a 
+practical static approach named \sys, to detect data races in OS kernels, and 
+estimate harmfulness of these data races. We have implemented \sys with 
+Clang~\cite{Clang} and Z3~\cite{Z3}. \sys automatically mines locking rules, 
+detects data races and estimates the harmfulness of detected data races. 
+Figure~\ref{fig_architecture} shows the architecture of \sys, which has four 
+phases:
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_architecture.pdf}
+	\figcaption{\sys architecture.}
+	\label{fig_architecture}
+\end{figure}
+
+\PP{P1: Source-code compilation.} The Clang compiler compiles the OS source 
+code into LLVM bytecode, and then the information collector scans each LLVM 
+bytecode file to record function information (including the position of each 
+function definition and function name, etc.) in a database. Such information is 
+used in subsequent code analysis for inter-procedural analysis across source 
+files.
+
+\PP{P2: Locking-rule Mining.} The locking-rule miner uses our alias-aware rule 
+mining method to deduce locking rules about whether a variable access should be 
+protected by a lock and which lock is required.
+
+\PP{P3: Data race detection and false-positive filtering.} The data race 
+detector detecting data races by checking whether a given access violate the 
+rules mined by our locking-rule miner. If so, the data race filter drops 
+false-positives with our locking-usage analysis. Besides, for a given data 
+race, there may be multiple code paths from the entry function to its 
+problematic instruction, and thus many repeated data races can be reported.  
+To drop repeated data races, for a new possible data race, the data race filter 
+checks whether its problematic instruction is identical to any existing data 
+race. If so, this possible data race is regarded as repeated and dropped.
+
+\PP{P4: Harmfulness estimation.} The harmfulness estimator uses our 
+pattern-based estimation to estimate the harmfulness of detected data races, 
+and reports data races that can cause memory or logic bugs. With these result, 
+developers can focus on those harmful data races.

### 2023-07-27 4.framework.tex
update evaluation

@@
 estimate harmfulness of these data races. We have implemented \sys with 
-Clang~\cite{Clang} and Z3~\cite{Z3}. \sys automatically mines locking rules, 
+Clang~\cite{clang} and Z3~\cite{z3}. \sys automatically mines locking rules, 
 detects data races and estimates the harmfulness of detected data races. 
@@
 \PP{P3: Data race detection and false-positive filtering.} The data race 
-detector detecting data races by checking whether a given access violate the 
-rules mined by our locking-rule miner. If so, the data race filter drops 
-false-positives with our locking-usage analysis. Besides, for a given data 
-race, there may be multiple code paths from the entry function to its 
-problematic instruction, and thus many repeated data races can be reported.  
-To drop repeated data races, for a new possible data race, the data race filter 
-checks whether its problematic instruction is identical to any existing data 
-race. If so, this possible data race is regarded as repeated and dropped.
+detector detecting data races by checking whether a given access violates the 
+rules mined by our locking-rule miner. After finding a possible data race, the 
+data race filter uses our locking-usage analysis to determine whether it is a 
+false positive. Besides, for a given data race, there may be multiple code 
+paths from the entry function to its problematic instruction, and thus many 
+repeated data races can be reported. To drop repeated data races, for a new 
+possible data race, the data race filter checks whether its problematic 
+instruction is identical to any existing data race. If so, this possible data 
+race is regarded as repeated and dropped.
 
@@
 pattern-based estimation to estimate the harmfulness of detected data races, 
-and reports data races that can cause memory or logic bugs. With these result, 
+and reports data races that can cause memory or logic bugs. With these results, 
 developers can focus on those harmful data races.

### 2023-07-27 5.evaluation.tex
update evaluation

@@
 \label{sec_evaluation}
+
+To validate the effectiveness of \sys, we evaluate it on the code of Linux 
+kernel 6.2. We run the evaluation on a regular x86-64 desktop with sixteen 
+Intel i7-10700 CPU@2.90GHz processors and 64GB physical memory. We use the 
+kernel configuration {\em allyesconfig} to enable all kernel code for the 
+x86-64 architecture.
+
+\begin{table}[tbph]
+	\tablecaption{Detection results of Linux 6.2.}
+	\label{tbl_bug_detection}
+	\renewcommand{\arraystretch}{1}
+	\setlength\tabcolsep{2pt}
+	\noindent{\scriptsize
+		\begin{center}
+			\begin{tabular}{p{1.5cm}|l|c}
+				\hline
+				\multicolumn{2}{c|}{\textbf{Description}} & \textbf{\sys}  
+				\\ \hline
+				\multirow{2}{1.5cm}{\textbf{{\em Code analysis}}} & 
+				Source files (analyzed/all) & xxxK/xxxK
+				\\ \cline{2-3}
+				& Source code lines (analyzed/all) & xxxM/xxxM 
+				\\ \cline{2-3}
+				\hline
+				\multirow{3}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
+				Key fields / total variables& 
+				\\ \cline{2-3}
+				& Mined locking rules &
+				\\ \cline{2-3}
+				& Protected field accesses / all accesses & 0.6
+				\\ \cline{2-3}
+				\hline
+				\multirow{2}{1.5cm}{\textbf{{\em Data race detection}}} & 
+				Detected data races (real / all) & 257 / 341
+				\\ \cline{2-3}
+				& Dropped data races by lock-usage analysis & 59
+				\\ \cline{2-3}
+				\hline
+				\multirow{5}{1.5cm}{\textbf{{\em Data race estimation}}}
+				& Null-pointer dereference (confirmed / all) & 15 / 20
+				\\ \cline{2-3}
+				& Infinite loop (confirmed / all) & 0 / 1
+				\\ \cline{2-3}
+				& Data inconsistency (confirmed / all) & 7 / 10
+				\\ \cline{2-3}
+				& Unprotected write (confirmed / all) & 10 / 57
+				\\ \cline{2-3}
+				& Total harmful data races (confirmed / all) & 32 / 88
+				\\ \cline{2-3}
+				\hline
+				\multirow{3}{1.5cm}{\textbf{{\em Time usage}}} & 
+				Key-field extraction & 
+				\\ \cline{2-3}
+				& Data-race detection &
+				\\ \cline{2-3}
+				& Total time &
+				\\ \cline{2-3}
+				\hline
+			\end{tabular}
+	\end{center}}
+\end{table}
+
+\subsection{Bug Detection}
+\label{subsec_bug_detection}
+
+We configure \sys with common lock-acquiring/release functions (like {\tt 
+spin\_lock and spin\_unlock}) to perform lock-set analysis to extract key 
+fields and mine locking rules, and lock-initialization functions (like {\tt 
+spin\_lock\_init}) to filter our false data races caused by code paths that 
+can not execute concurrently. And then run \sys to automatically check the 
+kernel source code. We manually check all the data races found by \sys, and 
+Table~\ref{tbl_bug_detection} shows the results, and source code lines are 
+counted by CLOC~\cite{cloc}. From the results, we have the following findings:
+
+\PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it in 
+total analyze xxxM lines of code in xxxK source code files within xxx hours. 
+The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
+are not enabled by the {\em allyesconfig} for the x86-64 architecture. We 
+believe that \sys can also find more data races in other architectures with 
+proper configuration.
+
+\PP{Locking-rule mining.} An OS kernel has a large code base with numerous 
+variables. Handling all variables when mining locking rules can introduce much 
+overhead. However, we observe that a variable tend to be protected by the lock 
+stored in the same data structure as the accessed variable. Based on this 
+observation, our locking-rule mining method first extract a key field by 
+finding whether there exists any access to it that is protected by a lock 
+stored in the same data structure. This method drops xxx\% variables (xxx out 
+xxx) that need to handled when mining locking rules, and thus can reduce 
+overhead significantly. After extracting key fields, our locking-rule mining 
+method collect all accessed to these key fields, and then deducing locking-rule 
+based on statistic. In this paper, given a data structure field, we set the 
+threshold of the ratio of accesses protected by a specific lock to all accesses 
+to 0.6. Our alias-aware rule mining method can drop many false rules and thus 
+only mines xxx rules, which can effectively reduce false data races.
+
+\PP{Data race detection.}
+
+\PP{Data race estimation.} Many data races are benign and can not cause memory 
+or logic bugs, and thus developers are unwilling to put effort into repairing 
+them. We exploit four patterns to detect null-pointer dereferences, infinite 
+loop, data inconsistency and unprotected write as introduced in 
+Section~\ref{subsec_estimation}, and find 88 data races in these patterns. We 
+report them to developers and 32 of them have been confirmed and fixed by them. 
+We still wait for response of other data races. Moreover, one of the bluetooth 
+l2cap protocol wonder if \sys can be used in their CI to detect these problems. 
+The results show that our pattern-based estimation can extract harmful data 
+races effectively, and can considerably reduce the workload of developers.

### 2023-07-28 5.evaluation.tex
udpate

@@
 				\multirow{2}{1.5cm}{\textbf{{\em Data race detection}}} & 
-				Detected data races (real / all) & 257 / 341
+				Detected data races (real / all) & 273 / 341
 				\\ \cline{2-3}
-				& Dropped data races by lock-usage analysis & 59
+				& Dropped data races by lock-usage analysis & 63
 				\\ \cline{2-3}
@@
 
-\PP{Data race detection.}
+\PP{Data race detection.} \sys reports 341 data races in the kernel source 
+code. We spent 15 hours on checking these data races and identify that 273 of 
+them are real, with a false positive rate of 19.9\%. Besides, our lock-usage 
+analysis drops 63 false data races and the results shows that it can reduce 
+false positives significantly.
 
@@
 report them to developers and 32 of them have been confirmed and fixed by them. 
-We still wait for response of other data races. Moreover, one of the bluetooth 
-l2cap protocol wonder if \sys can be used in their CI to detect these problems. 
+We still wait for response of other data races. Moreover, one of the developers 
+wonders if \sys can be used in their CI to detect these problems. 
 The results show that our pattern-based estimation can extract harmful data 
 races effectively, and can considerably reduce the workload of developers.
+
+\subsection{False Positives and Negatives}
+\label{subsec_false_pos_neg}
+\sys reports 68 false data races, and through manually checking these false 
+data races, we find that they are introduced for three main reasons:
+
+First, \sys employs an alias-aware rule mining method to deduce locking rules. 
+To improve the precision of mined rules, it assumes that the accessed variable 
+and the protecting lock are stored in the same data structure. Even so, \sys 
+can also deduce some false rules because some developers does not use locks 
+properly. They take a lock/unlock pair to protect accesses to all fields in the 
+same data structure, instead of the exact field should be protected for 
+convenience. As a result, our alias-aware rule mining method infers that 
+accesses to all fields surrounded by the lock/unlock pair should be protected 
+by the lock. This reason causes \sys to report 47 false data races.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_demo_false_rule.pdf}
+	\figcaption{A false data race caused by an incorrect locking rule.}
+	\label{fig_demo_false_rule}
+\end{figure}
+
+Figure~\ref{fig_demo_false_rule} shows a false data race caused by an incorrect 
+locking rule. In this example, only accesses to {\em cmd->t\_state} and {\em 
+cmd->transport\_state} should be protected by the lock {\em 
+cmd->t\_state\_lock}. However, the lock operation is put ahead of the switch 
+statement by developers, making our alias-aware rule mining method deduces that 
+{\em cmd->scsi\_status} and {\em cmd->se\_cmd\_flags} also need to be protected 
+by {\em cmd->t\_state\_lock} mistakenly. Based on this incorrect locking rule, 
+\sys reports a false positive at Line 745 when {\em cmd->se\_cmd\_flags} is 
+accessed in an if statement. Although this data race is a false positive, it 
+can cause performance degradation because the critical zone should have been 
+limited to Lines 899-900.
+
+Second, in order to reduce memory overhead and improve the performance of data 
+passing among different functions, an integer can be divided into several bit 
+vector to represent different data structure fields. However,

### 2023-07-29 3.technique.tex
update

@@
 
-At present, we propose four patterns that can cause null-pointer dereference, 
-infinite loop, data inconsistency and unprotected write, because these four 
-patterns are common and dangerous in OSes. The four patterns are shown in 
+At present, we propose three patterns that can cause null-pointer dereference, 
+data inconsistency and unprotected write, because these three patterns are 
+common and dangerous in OSes. The three patterns are shown in 
 Figure~\ref{fig_pattern}, and we assume that accesses to the fields of {\em 
@@
 	\centering
-	\includegraphics[width=0.85\linewidth]{figures/fig_pattern.pdf}
+	\includegraphics[width=1\linewidth]{figures/fig_pattern.pdf}
 	\figcaption{Three harmful patterns of data races.}
@@
 	and then dereferenced, a possible null-pointer dereference can occur.
-	\item \PP{Infinite loop.} The condition of the while statement does not 
-	protected by {\em dev->lock}. This can cause a infinite loop when {\em 
-	dev->cnt} is assigned with a non-zero value every time before it is checked 
-	by the while statement. The analysis detects such bugs by judge whether the 
-	data race acts as a condition of a loop statement.
+	%\item \PP{Infinite loop.} The condition of the while statement does not 
+	%protected by {\em dev->lock}. This can cause a infinite loop when {\em 
+	%dev->cnt} is assigned with a non-zero value every time before it is 
+	%checked 
+	%by the while statement. The analysis detects such bugs by judge whether 
+	%the 
+	%data race acts as a condition of a loop statement.
 	\item \PP{Data inconsistency.} Two different fields of the same data 

### 2023-07-29 5.evaluation.tex
update

@@
 				\hline
-				\multirow{5}{1.5cm}{\textbf{{\em Data race estimation}}}
+				\multirow{4}{1.5cm}{\textbf{{\em Data race estimation}}}
 				& Null-pointer dereference (confirmed / all) & 15 / 20
 				\\ \cline{2-3}
-				& Infinite loop (confirmed / all) & 0 / 1
-				\\ \cline{2-3}
+				%& Infinite loop (confirmed / all) & 0 / 1
+				%\\ \cline{2-3}
 				& Data inconsistency (confirmed / all) & 7 / 10
@@
 or logic bugs, and thus developers are unwilling to put effort into repairing 
-them. We exploit four patterns to detect null-pointer dereferences, infinite 
-loop, data inconsistency and unprotected write as introduced in 
+them. We exploit three patterns to detect null-pointer dereferences, data 
+inconsistency and unprotected write as introduced in 
 Section~\ref{subsec_estimation}, and find 88 data races in these patterns. We 
@@
 \label{subsec_false_pos_neg}
-\sys reports 68 false data races, and through manually checking these false 
-data races, we find that they are introduced for three main reasons:
+\PP{False positives.} \sys reports 68 false data races, and through manually 
+checking these false data races, we find that they are introduced for three 
+main reasons:
 
@@
 accessed in an if statement. Although this data race is a false positive, it 
-can cause performance degradation because the critical zone should have been 
-limited to Lines 899-900.
+can cause performance degradation because the critical zone protected by the 
+lock/unlock pair should have been limited to Lines 899-900.
 
@@
 passing among different functions, an integer can be divided into several bit 
-vector to represent different data structure fields. However,
+vectors to represent different data structure fields. However, in the LLVM 
+bytecode, accesses to these vectors are divided into a LOAD operation and 
+several bit operations, and thus \sys can not distinguish accesses to different 
+fields and reports false data races, because not all these fields should be 
+protected by a lock. This reason causes \sys to report 14 false data races.
+
+Apart from false data races caused by the two main reasons, there are still 7 
+other data races. \sys regards function without caller functions in a kernel 
+module as entry function and starts analysis from these entries. However, in 4 
+cases, an assertion is put at the beginning of the entry function, to guarantee 
+that the required lock is held, but \sys does not consider assertions and the 
+subsequent accesses are regarded as unprotected. Other 3 false data races are 
+accessed with {\em READ\_ONCE} or {\em WRITE\_ONCE}, and can not cause serious 
+issues.
+
+\PP{False negatives.} \sys may still miss some real data races for three main 
+reasons:
+
+First, our alias-aware mining method only considers accessed variables and 
+protecting locks that exist in the same data structure. However, on the one 
+hand, some accesses to variables may be protected by a global lock in a module, 
+and this lock is used to protect accesses to various variables and does 
+not exist in any data structure. On the other hand, some special lock 
+-acquiring/release functions (such as {\em rcu\_read\_lock} and {\em 
+rcu\_read\_unlock}) does not have an argument and is not handled in an alias 
+graph. Therefore, locking rules in the above two cases can be missed by our 
+alias-aware mining method. As a result, data races violated these locking rules 
+can not detected by \sys.
+
+Second, \sys does not handle function-pointer calls, and thus it can not build 
+complete call graphs for inter-procedural analysis. However, \sys starts 
+from functions without caller functions and performs analysis along call 
+graphs. And as a result, functions that are only called indirectly through 
+function pointers can not be analyzed by \sys, and thus data races involving 
+code reached through function-pointer calls can be missed.
+
+Finally, \sys exploits a lock-usage analysis to filter out false data races 
+caused by code paths that can not execute concurrently. Specifically, the 
+lock-usage analysis extracts all functions that are reachable from the lock 
+initialization functions in a call graph with the union-find set, and assumes 
+that these functions can not execute concurrently. However, some functions can 
+be called at multiple program sites, and some of the program sites may be not 
+reachable from lock initialization functions, but data races exist in these 
+functions are are all dropped by our lock-usage analysis.
+
+\subsection{Case Studies of the Found Harmful Data Races}
+\label{subsec_case_study}
+
+Figure~\ref{fig_case_bugs} shows three data races found by \sys in Linux 6.2,
+and they have been confirmed by Linux kernel developers.
+
+\begin{figure*}[htbp]
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fig_case_bugs.pdf}
+	\figcaption{Three real data races found by \sys in Linux 6.2.}
+	\label{fig_case_bugs}
+\end{figure*}
+
+\PP{Null-pointer dereference in the HDA sound driver.} In 
+Figure~\ref{fig_case_bugs}(a), the functions {\em snd\_hdac\_regmap\_exit()} 
+and {\em snd\_hdac\_regmap\_sync()} can execute concurrently. In Thread 2, the 
+variable {\em codec->regmap} is checked by an if statement in the function {\em 
+snd\_hdac\_regmap\_sync()}. Right after the condition of the if statement is 
+calculated to be true, the variable {\em codec->regmap} is set to be NULL by 
+the function {\em snd\_hdac\_regmap\_exit()} in Thread 2. And then the function 
+{\em regcache\_sync()} is called in Thread 1, with the argument {\em 
+codec->regmap}, after acquiring the lock {\em codec->regmap\_lock}. In the 
+called function, the variable {\em codec->regmap\_lock} is dereferenced through 
+{\em map->lock()}. In this execution case, the data structure field {\em 
+codec->regmap} is first assigned with NULL in Thread 1, and then dereferenced 
+in Thread 2, and thus cause a null-pointer dereference.
+
+\PP{Data inconsistency in the GFS2 file system.} In 
+Figure~\ref{fig_case_bugs}(b), the functions {\em gfs2\_show\_options()} and 
+{\em gfs2\_reconfigure()} can execute concurrently. In Thread 1, several fields 
+such as {\em gt\_statfs\_quantum} and {\em gt\_quota\_quantum} of {\em 
+sdp->sd\_tune} are accessed and their values are printed to logs. However, if 
+the value of {\em gt->gt\_quota\_quantum} is updated by the function {\em 
+gfs2\_reconfigure()} in Thread 2 right before the access to it in Thread 1, the 
+values of different fields recorded in logs can be inconsistent.
+
+\PP{Unprotected write in the LPFC SCSI driver.} In 
+Figure~\ref{fig_case_bugs}(c), the value of the shared variable {\em 
+phba->fcf.fcf\_flag} is set to 0 by the function {\em 
+lpfc\_unregister\_fcf\_rescan()}, without acquiring the lock {\em 
+phba->hbalock}, and this can cause double-fetch in several program sites. For 
+example, if {\em phba->fcf.fcf\_flag} is set to 0 by Thread 2 between the two 
+accesses to it in the function {\em lpfc\_register\_fcf()} in Thread 1 (Lines 
+1873 and 1874), the values of the two accesses can differ, causing 
+unpredictable behavior.

### 2023-07-31 1.introduction.tex
update

@@
 	a {\em lock-usage analysis} to filter out false positives caused by 
-	concurrency; (T3) a {\em pattern-based estimation} to extract harmful data 
-	races that can trigger memory or logical bugs such as null-pointer 
-	dereferences and data inconsistency.
+	code paths that can not execute concurrently; (T3) a {\em pattern-based 
+	estimation} to extract harmful data races that can trigger memory or 
+	logical bugs such as null-pointer dereferences and data inconsistency.	
 	\item Based on these three key techniques, we design a practical static 

### 2023-07-31 5.evaluation.tex
update

@@
 accesses to it in the function {\em lpfc\_register\_fcf()} in Thread 1 (Lines 
-1873 and 1874), the values of the two accesses can differ, causing 
-unpredictable behavior.
+1873 and 1874), the values of the two accesses can be different, causing 
+unpredictable behaviors.

### 2023-07-31 6.discussion.tex
update

@@
 \label{sec_discussion}
+
+\PP{Benefits in other aspects.} \sys uses an alias-aware mining method to 
+deduce precise locking rules about whether accesses to a variable should be 
+protected by a lock, and which lock are required. Such rules can benefit in 
+various aspects.
+
+First, with these rules, developers can improve reliability 
+and performance of kernel code. On the one hand, developers can carefully check 
+whether necessary lock are used when writing code, and avoid serious bugs 
+caused by data races in the early developing stage. On the other hand, 
+developers can limit the critical zone protected by an lock/unlock pair to the 
+exact accesses that should be protected, and let other code can execute 
+concurrently. And these locking rules can also help improve documents of OS 
+kernels, which is benefit for long term evolution of kernel codes.
+
+Second, key fields extract \sys are data structure fields that is protected by 
+some locks, and this means that these fields is likely to be concurrently 
+accessed by different threads. Therefore, key fields can be used in detecting 
+various types of concurrency bugs such as use-after-free and 
+uninitialized-variable access. Take concurrency use-after-free as an example, 
+if a key field is accessed in one thread and freed in another thread, a 
+concurrency use-after-free can occur.
+
+Finally, the locking rules deduced by our alias-aware rule mining method can 
+enhance other annotation-based analysis such as Clang thread safety 
+analysis~\cite{ClangThreadSafety}. Specifically, developers can use 
+instrumentation to annotate necessary locks for specific variables in the code 
+automatically, according to the locking rules generated by \sys, and run other 
+annotation-based analysis to detect data races.
+
+\PP{Comparison to other approaches.} We have try our best to experimentally 
+compare \sys to existing works in data race detection including 
+RELAY~\cite{Voung:FSE07}, WHOOP~\cite{Deligiannis:ASE15} and 
+Goblint~\cite{Vojdani:ASE16}. Unfortunately, they are too old and generate too 
+many errors when compiling and running, and thus we compare \sys to them 
+methodologically.
+
+\PP{Limitations and future works.} \sys can be improved in some aspects. First, 
+our alias-aware mining method only considers accessed variables and protecting 
+locks that exist in the same data structure, and can miss some locking rules 
+involves special locks such as RCU locks. To overcome this limitation, we plan 
+to create a common virtual node for RCU lock in the alias graph, and regard it 
+as the ancestor of all other nodes. In this way, all variables have a common 
+ancestor with the RCU lock, and can be inferred to be protected by the RCU 
+lock, because the accessed variable and the RCU lock are existing in the 
+identical data structure by our alias-aware rule mining method. Second, \sys 
+does not handle function-pointer calls, and thus it can not build complete call 
+graphs, and thus can miss data races reached from function-pointer calls. To 
+relieve this limitation, we plan to apply existing function-pointer 
+analysis~\cite{Hind:TOPLAS99, Zhang:ISPEC15, Heintze:PLDI01} in \sys, to detect 
+more data races in functions that called through function pointers and reduce 
+false negatives. Third, \sys performs a path-based analysis and suffers from 
+high time overhead, to accelerate its analysis, we plan to employ summer-based 
+analysis~\cite{Bai:ATC19, Bai:ATC22} to avoid re-analysis of the same function 
+definition. Finally, we plan to apply \sys to detect data races in other OS 
+kernels such as FreeBSD~\cite{FreeBSD} and NetBSD~\cite{NetBSD}, and use it to 
+detect more types of concurrency bugs including use-after-free and 
+uninitialized-variable access.

### 2023-07-31 7.related.tex
update

@@
 \label{sec_related}
+
+\subsection{Static Analysis of Data Races}
+\label{subsec_static}
+Many approaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, Anderson:PLDI09, 
+Zhou:MICRO19, Flanagan:PASTE01, Flanagan:PLDI00, Sadowski:PLATEAU14, 
+ClangThreadSafety, Blackshear:OOPSLA18, Choi:PLDI02, Engler:SOSP03, 
+Voung:FSE07, Pratikakis:PLDI06, Naik:PLDI06} use static analysis to detect data 
+races, without running the checked program. Some of 
+them~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, Anderson:PLDI09, Zhou:MICRO19, 
+Flanagan:PASTE01, Flanagan:PLDI00, Sadowski:PLATEAU14, ClangThreadSafety} 
+perform annotation-based analysis, and request developers to provide locking 
+rules about which lock is required when a specific variable is accessed, and 
+then detect data races that violate the provided rules. Clang thread safety 
+analysis~\cite{ClangThreadSafety} needs developers to annotate code with a {\tt 
+GUARDED\_BY} attribute to indicate which lock is required for a specific 
+variable, and then works like a type system for multi-threaded 
+programs~\cite{Vu:SoICT12, Boyapati:OOPSLA02, Marion:TAMC14} to detect data 
+races that violate the provided attribute. Flanagan et 
+al.~\cite{Flanagan:PLDI00} design a type system and relay on developers to 
+provide additional type annotations to associate a protecting lock with each 
+field declaration, and tracks the set of locks held at each program point. And 
+then it identifies whether a given program is race free by checking whether 
+necessary locks are indeed held at each program site. However, such 
+annotation-based approaches are not suitable for race detection in OS kernels, 
+because even an expert developer can not provide accurate annotations, because 
+the locking rules of kernel code are not well documented, and the logic of 
+kernel code tends to be very complex. Other 
+approaches~\cite{Blackshear:OOPSLA18, Choi:PLDI02, Engler:SOSP03, Voung:FSE07, 
+Pratikakis:PLDI06, Naik:PLDI06} employ lockset-based analysis to detect data 
+races automatically. They deduce locking rules based on statistic instead of 
+relying on manual annotations. But they does not consider alias 
+relationships~\cite{Voung:FSE07, Engler:SOSP03} or just employ imprecise 
+flow-insensitive alias analysis~\cite{Choi:PLDI02, 	Pratikakis:PLDI06, 
+Naik:PLDI06}. RacerX~\cite{Engler:SOSP03} performs a lockset 
+analysis to to detect data races in three modes, including simple checking, 
+simple statistical and precise statistical from least to most precise. However, 
+even if for the most precise mode, RacerX does not employ any alias analysis, 
+and thus introduce many false positives and negatives.
+
+Different from the above static analysis, \sys employs the alias graph to mine 
+locking rules automatically. Benefit from the path-based and field-sensitive 
+alias relationships generated by alias graph, \sys can deduce locking rules 
+effectively. With these locking rules, \sys can detect data races accurately. 
+Moreover, \sys exploits a pattern-based estimation to extract harmful data 
+races automatically, and can help developers focus on data races that can 
+indeed introduce serious issues.
+
+\subsection{Dynamic Analysis of Data Races}
+\label{subsec_dynamic}
+Runtime analysis does not suffer from the pointer alias problems, and thus many 
+approaches~\cite{Lochmann:EuroSys19, Lu:SOSP07, Lu:FSE18, Joshi:ASE08, 	
+Liu:NSDI07, Ocallahan:PPoPP03, Serebryany:WBIA09, Marino:PLDI09} detect data 
+races through dynamic analysis. LockDoc~\cite{Lochmann:EuroSys19} records 
+accesses to variables and lock acquisitions of an instrumented Linux kernel in 
+logs, and then infers locking rules from the logs. After that, LockDoc 
+automatically detect data races by checking whether a given variable access 
+violates the inferred locking rules. O'callahan et al.~\cite{Ocallahan:PPoPP03} 
+combine lockset-based detection and happens-before-based detection to detect 
+data races, which can get fewer false positives and less overhead than previous 
+dynamic analysis. In order to reduce overhead of runtime overhead further, 
+LiteRace~\cite{Marino:PLDI09} employs a sampling-based approach based on the 
+hypothesis that data races are likely to occur when a thread is executing
+an infrequently accessed region in the program. It instruments sampled memory 
+accesses, and perform happens-before-based analysis to detect data races, 
+according the logs generated by the instrumentation.
+
+However, dynamic analysis suffers from low code coverage, and thus the locking 
+rules inferred through logs can be imprecise. Besides, dynamic analyses need to 
+run OS kernels with instrumentation and are hard to deploy.

### 2023-07-31 8.conclusion.tex
update

@@
 \label{sec_conclusion}
+
+In this paper, we develop a novel static approach named \sys to detect data 
+races in OS kernels by mining locking rules. It consists of three key 
+techniques, including an alias-aware rule mining method to automatically deduce 
+locking rules, a lock-usage analysis to filter out false positives caused by 
+code paths that can not execute concurrently and a pattern-based estimation to 
+extract harmful data races that can trigger memory or logical bugs such as 
+null-pointer dereferences and data inconsistency. In the evaluation, \sys finds 
+88 real harmful data races, and 32 of them have been confirmed by the 
+developers.

### 2023-08-01 2.motivation.tex
Finish the draft of technique
udpate
Fix some grammar issues

@@
 % !TeX spellcheck = en_US
-\section{Background and Motivation}
+\section{Motivation}
 \label{sec_motivation}
@@
 \end{figure}
+
 Figure~\ref{fig_demo_bug} shows a real null-pointer dereference caused by a 
@@
 can execute concurrently. We exploit \textcircled{\footnotesize{n}} to 
-represent the execution order of instructions and show one execution case in 
+represent the execution order of instructions and show one execution case on 
 the left of Figure~\ref{fig_demo_bug}. In Thread 1, the variables {\tt 
-crtc->state->event} and {\tt crtc->state->active} is checked by an if statement 
-in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. After the condition 
-is calculated to be true, the function {\tt exynos\_crtc\_handle\_event()} is 
-executed in Thread 2. In this function, the value of {\tt crtc->state->active} 
-is assigned to {\tt event} (\textcircled{\footnotesize{2}}) and then {\tt 
-event} is checked in an if statement (\textcircled{\footnotesize{3}}). If it is 
-not NULL, the variable {\tt crtc->state->event} is assigned with NULL 
+crtc->state->event} and {\tt crtc->state->active} are checked by an if 
+statement in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. After the 
+condition is calculated to be true, the function {\tt 
+exynos\_crtc\_handle\_event()} is executed in Thread 2. In this function, the 
+value of {\tt crtc->state->active} is assigned to {\tt event} 
+(\textcircled{\footnotesize{2}}) and then {\tt event} is checked in an if 
+statement (\textcircled{\footnotesize{3}}). If it is not NULL, the variable 
+{\tt crtc->state->event} is assigned with NULL 
 (\textcircled{\footnotesize{4}}). Right after this assignment, the function 
@@
 Thread 2 right after the first condition of the if statement in Thread 1 is 
-calculated to be true. Such requirement is difficult to satisfy by executing 
+calculated to be true. Such a requirement is difficult to satisfy by executing 
 existing test suites. In fact, this bug had existed for nearly 6 years since 
@@
 \PP{C1: Getting locking rules.} The relationship between variables and locks 
-are not well documented in OS kernels, making it hard to determine whether a 
+is not well documented in OS kernels, making it hard to determine whether a 
 specific variable should be protected by a lock and which lock is 
@@
 \PP{C2: Dropping false data races.} Static analysis suffers from false 
-positives. For example, each data race involves more than one code paths that 
+positives. For example, each data race involves more than one code path that 
 should be able to concurrently execute. However, which code can execute 
@@
 approaches~\cite{Narayanasamy:PLDI07, Sen:PLDI08, Kasikci:SOSP13, 
-Kasikci:ASPLOS12} estimate harmfulness of data races through dynamic analysis, 
-and explore thread interleavings to trigger data races to estimate their 
-harmfulness. However, they suffer from low code coverage and thus can miss many 
-real harmful data races.
+Kasikci:ASPLOS12} estimate the harmfulness of data races through dynamic 
+analysis, and explore thread interleavings to trigger data races to estimate 
+their harmfulness. However, they suffer from low code coverage and thus can 
+miss many real harmful data races.

### 2023-08-01 3.technique.tex
Fix some grammar issues

@@
 locking rules. For {\em C2}, we propose a {\em lock-usage analysis} to filter 
-out false data races by validating concurrency of code paths. For {\em C3}, we 
-propose a pattern-based estimation to extract harmful races that can trigger 
-memory or logical bugs such as null-pointer dereference and data inconsistency. 
-We introduce them as follows:
+out false data races by validating the concurrency of kernel code. For {\em 
+C3}, we propose a pattern-based estimation to extract harmful races that can 
+trigger memory or logical bugs such as null-pointer dereference and data 
+inconsistency. We introduce them as follows:
 
@@
 \label{subsec_rule_mining}
-The relationship between variables and locks are not well documented in OS 
+The relationship between variables and locks is not well documented in OS 
 kernels, but it can be inferred from the kernel code. Specifically, a variable 
@@
 in most cases, it is likely to be protected by the lock. Whether a variable and 
-the protecting lock exist in the same data structure can be determined though 
+the protecting lock exist in the same data structure can be determined through 
 an alias graph~\cite{Li:ASPLOS22, Kastrinis:CC18} by finding their common 
@@
 
-\PP{Alias Graph.} It is an important data structure to infer relationships 
+\PP{Alias Graph.} It is an important data structure to infer relationship 
 between variable and its protecting lock in our analysis, so we introduce it 
@@
 alias node, a variable residing in a node followed by a sequence of edge
-labels form an access path~\cite{Kastrinis:CC18, Cheng:PLDI00}. In this paper, 
+labels forms an access path~\cite{Kastrinis:CC18, Cheng:PLDI00}. In this paper, 
 we replace the variable in an access path with its structure name to represent 
@@
 the destination node of the edge that comes from $\mathit{n_2}$ and is labeled 
-with $\mathit{*}$, and then move $\mathit{v_1}$ to the destination node. And 
+with $\mathit{*}$, and then moves $\mathit{v_1}$ to the destination node. And 
 after this operation, $\mathit{v_1}$ and $\mathit{*v_2}$ are represented by the 
@@
 \noindent{\textbf{\em Example.}} Figure~\ref{fig_demo_alias_graph} shows a 
-piece of driver-like source code and its alias graph. In this example, after an 
+piece of driver-like source code and its alias graph. In this example, after a 
 GEP (\&dev->lock) operation at Line 6, an edge labeled with {\tt lock} from 
@@
 structure (namely {\tt Dev}). Therefore, the structure field Dev.flag is likely 
-to be protected the lock stored in the structure field Dev.lock.
+to be protected by the lock stored in the structure field Dev.lock.
 
@@
 alias-aware rule mining method has three main stages shown in 
-Figure~\ref{fig_workflow}. In Stage 1, it extract key variable that may be 
-protected by a lock. In Stage 2, it collect all accesses to key fields as well 
-as code paths these accesses exist in. In Stage 3, it deduce locking rules by 
-calculate the proportion of field accesses protected by a specific lock in all 
-field accesses.
+Figure~\ref{fig_workflow}. In Stage 1, it extracts key variables that may be 
+protected by a lock. In Stage 2, it collects all accesses to key fields as well 
+as code paths these accesses exist in. In Stage 3, it deduces locking rules by 
+calculating the proportion of field accesses protected by a specific lock in 
+all field accesses.
 
@@
 much unnecessary overhead. Generally, only a few field accesses can miss 
-necessary protecting lock, due to carelessness of developers. Based on this 
-insight, our analysis first extracts key fields that may need to be protected 
-by a specific lock, by performing a lock-set analysis to find whether a given 
-variable is accessed after acquiring a lock in the same data structure as the 
-accessed field in any code path.
+necessary protecting locks, due to the carelessness of developers. Based on 
+this insight, our analysis first extracts key fields that may need to be 
+protected by a specific lock, by performing a lock-set analysis to find whether 
+a given variable is accessed after acquiring a lock in the same data structure 
+as the accessed field in any code path.
 
@@
 graph. The analysis first gets the operand of the instruction (Line 1), and 
-then get the node of the operand from the alias graph (Line 2). If the 
+then gets the node of the operand from the alias graph (Line 2). If the 
 instruction is a lock operation, the node of the operand is inserted into the 
@@
 the instruction is a write or a read, the analysis first gets the node of the 
-operand with the new alias graph (Lines 4-5). And then, for each node of 
+operand with the new alias graph (Lines 4-5). And then, for each node of the 
 lock variable in the lock set, the analysis uses the alias graph to extract the 
@@
 inserted into the set of key fields (Lines 9-11). For inter-procedural 
-analysis, the analysis copies function definition into its call sites and this 
+analysis, the analysis copies function definition into its call sites, but this 
 operation is not shown in Figure~\ref{fig_pseudocode_field_extract} for 
@@
 \PP{S2: Field-access collection.} With the key fields extracted in Stage 1, the 
-analysis can only collects accesses to the variables that is protected in some 
-cases, because if a variable is never protected by any lock when is accessed, 
-it is less likely to be shared by different threads, and thus can not introduce 
-no concurrency issue.
+analysis only needs to collect accesses to the variables that are protected in 
+some cases, because if a variable is never protected by any lock when is 
+accessed, it is less likely to be shared by different threads, and thus can not 
+introduce any concurrency issue.
 
@@
 write or a read) is set according to the instruction (Lines 5-8). However, if 
-the instruction is not an access operation, the function returns with a NULL 
-value for the field access (Line 10). Otherwise, the analysis first gets the 
-node of the operand with the new alias graph (Lines 12-13), and then for each 
-node of the lock variable in the lock set, the analysis uses the alias graph 
-toextract the fields that the accessed variable and the lock variable stored in 
-(Lines 16-17). If the fields are found, the field access is returned with the 
-found fields (Lines 18-22). If the fields are not found, the accessed variable  
-does not protected by any lock. In this case, if the accessed variable is 
-stored in a key field, the field access is returned with a NULL lock variable 
-(Lines 25-30).
+the instruction is not an access operation, the function returns a NULL value 
+for the field access (Line 10). Otherwise, the analysis first gets the node of 
+the operand with the new alias graph (Lines 12-13), and then for each node of 
+the lock variable in the lock set, the analysis uses the alias graph to extract 
+the fields that the accessed variable and the lock variable stored in (Lines 
+16-17). If the fields are found, the field access is returned with the found 
+fields (Lines 18-22). If the fields are not found, the accessed variable  are 
+not protected by any lock. In this case, if the accessed variable is stored in 
+a key field, the field access is returned with a NULL lock variable (Lines 
+25-30).
 
@@
 statement and cause numerous code paths for the same variable access. Based on 
-these consideration, the analysis distinguishes different accesses to the same 
+this consideration, the analysis distinguishes different accesses to the same 
 data structure field by different calling contexts. Specifically, given a key 
@@
 ratio {\em locked\_access\_num}/{\em all\_access\_num} is larger than a given 
-threshold, and these is at least one write access to {\em f}, the data 
+threshold, and there is at least one write access to {\em f}, the data 
 structure field {\em f} is inferred to be protected by the lock {\em l}. After 
-deducting locking rules, the analysis detecting data races by checking whether 
-a field access validates any locking rules.
+deducting locking rules, the analysis detects data races by checking whether a 
+field access validates any locking rules.
 
@@
 Figure~\ref{fig_demo_rule_mining} to illustrate how to mine locking rules 
-with the three stages. The analysis first perform a path-based analysis to 
+with the three stages. The analysis first performs a path-based analysis to 
 extract key fields. Take the code path {\tt Line 12, Line 6, Line 7, Line 8, 
@@
 ({\em *fp = f}). The final alias graph is shown in 
-Figure~\ref{fig_demo_rule_mining}(b). At the same time of alias analysis, the 
+Figure~\ref{fig_demo_rule_mining}(b). At the same time as alias analysis, the 
 analysis records the acquired lock {\em Dev.lock} into the lock set after the 
@@
 two of them are protected by the lock {\em Dev.lock}. And thus the two fields 
-{\em Dev.flag} and {\em Dev.flag*} is deduce to be protected by the lock {\em 
+{\em Dev.flag} and {\em Dev.flag*} are deduced to be protected by the lock {\em 
 Dev.lock} (assume the threshold of the ratio {\em locked\_access\_num} / 
@@
 \label{subsec_lock_usage_analysis}
-Our analysis takes functions that has no caller function as the entry as 
+Our analysis takes functions that have no caller function as the entry as 
 existing work~\cite{Li:ASPLOS22}, and performs a path-based analysis. It 
@@
 concurrently in fact. We observe that each kernel module has an initialization 
-phase, after which many function can be called concurrently by other modules. 
+phase, after which many functions can be called concurrently by other modules. 
 When performing initialization, the kernel module serially initialized the lock 
@@
 Our lock-usage analysis uses the union-find set~\cite{Galler:ACM64} to get all 
-functions that reachable from the lock initialization function such as {\em 
+functions that are reachable from the lock initialization function such as {\em 
 mutex\_init()} and {\em spin\_lock\_init()}. 
@@
 analysis. For each analyzed function in the OS kernel, the analysis first gets 
-all the called function of it (Lines 3-4), and then union the analyzed function 
-and the called function to update the union-find set (Line 5). After processing 
-all the call instructions, the analysis judging whether a function executes in 
-the initialization phase by checking whether it exist in the same set with the 
-given lock initialization function (Lines 9-11).
+all the called functions of it (Lines 3-4), and then union the analyzed  
+function and the called function to update the union-find set (Line 5). After 
+processing all the call instructions, the analysis judges whether a function 
+executes in the initialization phase by checking whether it exists in the same 
+set with the given lock initialization function (Lines 9-11).
 
@@
 inferred to execute in the module initialization phase and can not execute 
-concurrently with other function, and this is a false data race.
+concurrently with other functions, and this is a false data race.
 
@@
 Many data races are benign or introduced by developers deliberately to improve 
-the kernel performance. They can not cause memory or logic bugs, and thus 
+kernel performance. They can not cause memory or logic bugs, and thus 
 developers are unwilling to put effort into repairing them. We observe that 
@@
 	dereference if {\em dev->event} is set to NULL right after the condition of 
-	the if statement is checked to be true. To recognize such pattern, the 
+	the if statement is checked to be true. To recognize such a pattern, the 
 	analysis first locates the data race, if it is checked by an if statement 
@@
 	changed by another thread right after the access to {\em dev->clock->min}. 
-	To recognize such pattern, the analysis detect whether multiple fields of 
-	the same data structure are accessed without acquiring the protecting lock.
+	To recognize such a pattern, the analysis detects whether multiple fields 
+	of 	the same data structure are accessed without acquiring the protecting 
+	lock.
 	\item \PP{Unprotected write.} The write access to {\em dev->event} is not 
@@
 	dev->event} can be modified at any time when other threads access it, and 
-	thus introduce unpredictable behavior. The analysis detect such pattern by 
-	judge whether a data race occurs in a write access.
+	thus introduce unpredictable behavior. The analysis detects such a pattern 
+	by judging whether a data race occurs in a write access.
 \end{itemize}

### 2023-08-01 4.framework.tex
Fix some grammar issues

@@
 practical static approach named \sys, to detect data races in OS kernels, and 
-estimate harmfulness of these data races. We have implemented \sys with 
+estimate the harmfulness of these data races. We have implemented \sys with 
 Clang~\cite{clang} and Z3~\cite{z3}. \sys automatically mines locking rules, 
@@
 \PP{P3: Data race detection and false-positive filtering.} The data race 
-detector detecting data races by checking whether a given access violates the 
+detector detects data races by checking whether a given access violates the 
 rules mined by our locking-rule miner. After finding a possible data race, the 

### 2023-08-01 5.evaluation.tex
Fix some grammar issues

@@
 \PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it in 
-total analyze xxxM lines of code in xxxK source code files within xxx hours. 
+total analyzes xxxM lines of code in xxxK source code files within xxx hours. 
 The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
@@
 variables. Handling all variables when mining locking rules can introduce much 
-overhead. However, we observe that a variable tend to be protected by the lock 
+overhead. However, we observe that a variable tends to be protected by the lock 
 stored in the same data structure as the accessed variable. Based on this 
-observation, our locking-rule mining method first extract a key field by 
+observation, our locking-rule mining method first extracts a key field by 
 finding whether there exists any access to it that is protected by a lock 
 stored in the same data structure. This method drops xxx\% variables (xxx out 
-xxx) that need to handled when mining locking rules, and thus can reduce 
+xxx) that need to be handled when mining locking rules, and thus can reduce 
 overhead significantly. After extracting key fields, our locking-rule mining 
-method collect all accessed to these key fields, and then deducing locking-rule 
-based on statistic. In this paper, given a data structure field, we set the 
+method collects all accesses to these key fields, and then deduces locking-rule 
+based on statistics. In this paper, given a data structure field, we set the 
 threshold of the ratio of accesses protected by a specific lock to all accesses 
@@
 them are real, with a false positive rate of 19.9\%. Besides, our lock-usage 
-analysis drops 63 false data races and the results shows that it can reduce 
+analysis drops 63 false data races and the results show that it can reduce 
 false positives significantly.
@@
 report them to developers and 32 of them have been confirmed and fixed by them. 
-We still wait for response of other data races. Moreover, one of the developers 
-wonders if \sys can be used in their CI to detect these problems. 
+We still wait for the response of other data races. Moreover, one of the 
+developers wonders if \sys can be used in their CI to detect these problems. 
 The results show that our pattern-based estimation can extract harmful data 
@@
 and the protecting lock are stored in the same data structure. Even so, \sys 
-can also deduce some false rules because some developers does not use locks 
+can also deduce some false rules because some developers do not use locks 
 properly. They take a lock/unlock pair to protect accesses to all fields in the 
-same data structure, instead of the exact field should be protected for 
+same data structure, instead of the exact field that should be protected for 
 convenience. As a result, our alias-aware rule mining method infers that 
@@
 cmd->t\_state\_lock}. However, the lock operation is put ahead of the switch 
-statement by developers, making our alias-aware rule mining method deduces that 
+statement by developers, making our alias-aware rule mining method deduce that 
 {\em cmd->scsi\_status} and {\em cmd->se\_cmd\_flags} also need to be protected 
@@
 Apart from false data races caused by the two main reasons, there are still 7 
-other data races. \sys regards function without caller functions in a kernel 
-module as entry function and starts analysis from these entries. However, in 4 
-cases, an assertion is put at the beginning of the entry function, to guarantee 
-that the required lock is held, but \sys does not consider assertions and the 
-subsequent accesses are regarded as unprotected. Other 3 false data races are 
-accessed with {\em READ\_ONCE} or {\em WRITE\_ONCE}, and can not cause serious 
-issues.
+other data races. \sys regards a function without caller functions in a kernel 
+module as the entry function and starts analysis from these entries. However, 
+in 4 cases, an assertion is put at the beginning of the entry function, to 
+guarantee that the required lock is held, but \sys does not consider assertions 
+and the subsequent accesses are regarded as unprotected. The other 3 false data 
+races are accessed with {\em READ\_ONCE} or {\em WRITE\_ONCE}, and can not 
+cause serious issues.
 
@@
 -acquiring/release functions (such as {\em rcu\_read\_lock} and {\em 
-rcu\_read\_unlock}) does not have an argument and is not handled in an alias 
+rcu\_read\_unlock}) do not have an argument and are not handled in an alias 
 graph. Therefore, locking rules in the above two cases can be missed by our 
-alias-aware mining method. As a result, data races violated these locking rules 
-can not detected by \sys.
+alias-aware mining method. As a result, data races that violate these locking 
+rules can not be detected by \sys.
 
@@
 be called at multiple program sites, and some of the program sites may be not 
-reachable from lock initialization functions, but data races exist in these 
-functions are are all dropped by our lock-usage analysis.
+reachable from the lock initialization functions, but data races that exist in 
+these functions are all dropped by our lock-usage analysis.
 

### 2023-08-01 6.discussion.tex
Fix some grammar issues

@@
 deduce precise locking rules about whether accesses to a variable should be 
-protected by a lock, and which lock are required. Such rules can benefit in 
+protected by a lock, and which lock is required. Such rules can benefit in 
 various aspects.
 
-First, with these rules, developers can improve reliability 
+First, with these rules, developers can improve the reliability 
 and performance of kernel code. On the one hand, developers can carefully check 
-whether necessary lock are used when writing code, and avoid serious bugs 
+whether necessary locks are used when writing code, and avoid serious bugs 
 caused by data races in the early developing stage. On the other hand, 
-developers can limit the critical zone protected by an lock/unlock pair to the 
+developers can limit the critical zone protected by a lock/unlock pair to the 
 exact accesses that should be protected, and let other code can execute 
 concurrently. And these locking rules can also help improve documents of OS 
-kernels, which is benefit for long term evolution of kernel codes.
+kernels, which is beneficial  for long term evolution of kernel codes.
 
-Second, key fields extract \sys are data structure fields that is protected by 
-some locks, and this means that these fields is likely to be concurrently 
-accessed by different threads. Therefore, key fields can be used in detecting 
-various types of concurrency bugs such as use-after-free and 
+Second, key fields extracted by \sys are data structure fields that are 
+protected by some locks, and this means that these fields are likely to be 
+concurrently accessed by different threads. Therefore, key fields can be used 
+in detecting various types of concurrency bugs such as use-after-free and 
 uninitialized-variable access. Take concurrency use-after-free as an example, 
@@
 Finally, the locking rules deduced by our alias-aware rule mining method can 
-enhance other annotation-based analysis such as Clang thread safety 
+enhance other annotation-based analyses such as Clang thread safety 
 analysis~\cite{ClangThreadSafety}. Specifically, developers can use 
@@
 
-\PP{Comparison to other approaches.} We have try our best to experimentally 
+\PP{Comparison to other approaches.} We have tried our best to experimentally 
 compare \sys to existing works in data race detection including 
@@
 locks that exist in the same data structure, and can miss some locking rules 
-involves special locks such as RCU locks. To overcome this limitation, we plan 
+involving special locks such as RCU locks. To overcome this limitation, we plan 
 to create a common virtual node for RCU lock in the alias graph, and regard it 
@@
 analysis~\cite{Hind:TOPLAS99, Zhang:ISPEC15, Heintze:PLDI01} in \sys, to detect 
-more data races in functions that called through function pointers and reduce 
-false negatives. Third, \sys performs a path-based analysis and suffers from 
-high time overhead, to accelerate its analysis, we plan to employ summer-based 
-analysis~\cite{Bai:ATC19, Bai:ATC22} to avoid re-analysis of the same function 
-definition. Finally, we plan to apply \sys to detect data races in other OS 
-kernels such as FreeBSD~\cite{FreeBSD} and NetBSD~\cite{NetBSD}, and use it to 
-detect more types of concurrency bugs including use-after-free and 
-uninitialized-variable access.
+more data races in functions that are called through function pointers and 
+reduce false negatives. Third, \sys performs a path-based analysis and suffers 
+from high time overhead, to accelerate its analysis, we plan to employ 
+summer-based analysis~\cite{Bai:ATC19, Bai:ATC22} to avoid re-analysis of the 
+same function definition. Finally, we plan to apply \sys to detect data races 
+in other OS kernels such as FreeBSD~\cite{FreeBSD} and NetBSD~\cite{NetBSD}, 
+and use it to detect more types of concurrency bugs including use-after-free 
+and uninitialized-variable access.

### 2023-08-01 7.related.tex
Fix some grammar issues

@@
 races that violate the provided attribute. Flanagan et 
-al.~\cite{Flanagan:PLDI00} design a type system and relay on developers to 
+al.~\cite{Flanagan:PLDI00} design a type system and rely on developers to 
 provide additional type annotations to associate a protecting lock with each 
-field declaration, and tracks the set of locks held at each program point. And 
+field declaration, and track the set of locks held at each program point. And 
 then it identifies whether a given program is race free by checking whether 
@@
 Pratikakis:PLDI06, Naik:PLDI06} employ lockset-based analysis to detect data 
-races automatically. They deduce locking rules based on statistic instead of 
-relying on manual annotations. But they does not consider alias 
+races automatically. They deduce locking rules based on statistics instead of 
+relying on manual annotations. But they do not consider alias 
 relationships~\cite{Voung:FSE07, Engler:SOSP03} or just employ imprecise 
 flow-insensitive alias analysis~\cite{Choi:PLDI02, 	Pratikakis:PLDI06, 
-Naik:PLDI06}. RacerX~\cite{Engler:SOSP03} performs a lockset 
-analysis to to detect data races in three modes, including simple checking, 
-simple statistical and precise statistical from least to most precise. However, 
-even if for the most precise mode, RacerX does not employ any alias analysis, 
-and thus introduce many false positives and negatives.
+Naik:PLDI06}. RacerX~\cite{Engler:SOSP03} performs a lockset analysis to detect 
+data races in three modes, including simple checking, simple statistical and 
+precise statistical from least to most precise. However, even if for the most 
+precise mode, RacerX does not employ any alias analysis, and thus introduces 
+many false positives and negatives.
 
 Different from the above static analysis, \sys employs the alias graph to mine 
-locking rules automatically. Benefit from the path-based and field-sensitive 
-alias relationships generated by alias graph, \sys can deduce locking rules 
+locking rules automatically. Benefiting from the path-based and field-sensitive 
+alias relationships generated by alias graphs, \sys can deduce locking rules 
 effectively. With these locking rules, \sys can detect data races accurately. 
@@
 \label{subsec_dynamic}
-Runtime analysis does not suffer from the pointer alias problems, and thus many 
-approaches~\cite{Lochmann:EuroSys19, Lu:SOSP07, Lu:FSE18, Joshi:ASE08, 	
-Liu:NSDI07, Ocallahan:PPoPP03, Serebryany:WBIA09, Marino:PLDI09} detect data 
+The runtime analysis does not suffer from the pointer alias problems, and thus 
+many approaches~\cite{Lochmann:EuroSys19, Lu:SOSP07, Lu:FSE18, Joshi:ASE08, 	
+Liu:NSDI07, Ocallahan:PPoPP03, Serebryany:WBIA09, Marino:PLDI09} detects data 
 races through dynamic analysis. LockDoc~\cite{Lochmann:EuroSys19} records 
@@
 data races, which can get fewer false positives and less overhead than previous 
-dynamic analysis. In order to reduce overhead of runtime overhead further, 
+dynamic analyses. In order to reduce runtime overhead further, 
 LiteRace~\cite{Marino:PLDI09} employs a sampling-based approach based on the 
@@
 an infrequently accessed region in the program. It instruments sampled memory 
-accesses, and perform happens-before-based analysis to detect data races, 
-according the logs generated by the instrumentation.
+accesses, and performs happens-before-based analysis to detect data races, 
+according to the logs generated by the instrumentation.
 

### 2023-08-02 1.introduction.tex
Finish a temporary version (without comparing experiment and some evaluation results)

--- 2023-07-23 1.introduction.tex
@@
 introduce concurrency issues. Data race is one of the most harmful concurrency 
-issues that occurs when multiple threads access the same variable concurrently 
-and at least one of the accesses is a write. Some data races are quite 
-dangerous and can cause serious bugs such as null-pointer dereferences, data 
-inconsistency and infinite loops.
+issues that occurs when multiple threads access a shared variable concurrently 
+and at least one of the accesses is a write. Some races are harmful and can 
+cause serious bugs like null-pointer dereferences, data inconsistencies and 
+double fetches.
 
@@
 rely on developers to supply annotations that describe the locking discipline 
-such as which lock protects a specific variable, and detect data races by 
-exploiting variable accesses that violate these disciplines. For example, Clang 
-thread safety analysis~\cite{ClangThreadSafety} requires developers to label a 
-variable and the lock to protect it with a {\tt GUARDED\_BY} attribute. And 
-then detects data races by finding variable accesses that violate the given 
-attribute. However, such annotation-based approaches are not suitable for race 
-detection in OS kernels, because even an expert developer can not provide 
-accurate annotations, due to poor documentation and complicated logic of OS 
-codes.
+such as which lock is required for accesses to a specific variable, and detect 
+data races by searching variable accesses that violate these disciplines. For 
+example, Clang thread safety analysis~\cite{ClangThreadSafety} requires 
+developers to label a variable and the lock to protect it with a {\tt 
+GUARDED\_BY} attribute. And then detects data races by finding variable 
+accesses that violate the given attribute. However, such annotation-based 
+approaches are not suitable for race detection in OS kernels, because even an 
+expert developer can not provide accurate annotations, due to poor document and 
+complicated logic of OS codes.
 
-Some static approaches~\cite{Choi:PLDI02, Engler:SOSP03, Voung:FSE07, 
+Other static approaches~\cite{Choi:PLDI02, Engler:SOSP03, Voung:FSE07, 
 Pratikakis:PLDI06, Naik:PLDI06} employ lockset-based analysis to detect data 
-races automatically. They first collect locks that are acquired when a given 
-variable is accessed, and then mine locking rules about whether a given 
-variable should be protected by a specific lock. At last, they detect data 
-races by checking whether variable accesses violate the mined locking rules. 
-However, they do not consider alias relations~\cite{Voung:FSE07, Engler:SOSP03} 
-or just use imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, 
-Pratikakis:PLDI06, Naik:PLDI06}, and thus can introduce both false positives 
-and false negatives.
+races automatically. For a given variable, they first collect locks that are 
+acquired for each access to it, and then detect data races by checking whether 
+the intersection of locks for different accesses is empty. However, they do not 
+consider alias relationships~\cite{Voung:FSE07, Engler:SOSP03} or just use 
+imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, Pratikakis:PLDI06, 
+Naik:PLDI06}, and thus can introduce both false positives and negatives.
 
@@
 acquisitions of an instrumented Linux kernel, and then infers locking rules 
-from the recorded accesses. After that, LockDoc automatically locates variables 
-access that violate the inferred locking rules to detect concurrency bugs such 
-as data races. However, dynamic analysis suffers from low code coverage, and 
-thus the locking rules inferred through execution traces can be imprecise. 
+from the recorded accesses. After that, LockDoc automatically locates variable 
+accesses that violate the inferred locking rules to detect concurrency issues 
+such as data races. However, dynamic analysis suffers from low code coverage, 
+and thus the locking rules inferred through execution traces can be imprecise. 
 Besides, dynamic analyses are hard to deploy, because they need to run OS 
@@
 to automatically detect data races and estimate the harmfulness of these data 
-races in OS kernels effectively. \sys consists of three key techniques:
+races in OS kernels. \sys consists of three key techniques:
 
-(T1) \sys uses an {\em alias-aware rule mining method} to automatically 
-deduce locking rules. We observe that an access to a variable is often 
-protected by the lock in the same data structure as the accessed field. And 
-this relationship between accessed field and the protecting lock can be 
-effectively inferred from a special data structure named alias 
-graph~\cite{Li:ASPLOS22, Kastrinis:CC18}. Specifically, fields in the same data 
-structure have the same ancestor in the alias graph, and thus whether an 
-accessed field and a specific lock are in the same data structure can be 
-inferred by finding a common ancestor in the alias graph. Moreover, with 
-benefits from precise field-sensitive alias relationships of alias graph, 
-\sys can effectively extract the accessed field and its corresponding 
-protecting lock. After obtaining the accessed field and protecting lock pairs, 
-\sys calculates the proportion of field accesses protected by a specific 
-lock in all field accesses. If the proportion is larger than a given threshold, 
-the accessed field is determined to be protected by the protecting lock.
+(T1) \sys uses an {\em alias-aware rule mining method} to automatically deduce 
+locking rules about whether accesses to a specific data structure field should 
+be protected and which data structure field the protecting lock exist in. We 
+observe that the variable that is accessed and the lock to protect the access 
+often exist in the same data structure. And this relationship between the 
+accessed variable and the protecting lock can be effectively inferred from a 
+special data structure named alias graph~\cite{Li:ASPLOS22, Kastrinis:CC18}. 
+Specifically, variables in the same data structure have a common ancestor in 
+the alias graph, and thus whether an accessed variable and a specific lock are 
+in the same data structure can be inferred through finding a common ancestor in 
+the alias graph. Moreover, with benefits from precise field-sensitive alias 
+relationships of alias graph, \sys can effectively extract the accessed field 
+and its corresponding lock field. After obtaining the accessed field and lock 
+field pairs, \sys calculates the proportion of field accesses protected by a 
+specific lock field in all field accesses. If the proportion is larger than a 
+given threshold, the accessed field is determined to be protected by the lock 
+field for a specific data structure.
 
-(T2) \sys uses a {\em lock-usage analysis} to filter out false positives 
-validating concurrency of code paths. After mining locking rules, \sys detects 
-data races by checking whether a given variable access violates the 
-locking rules. To reduce false negatives, \sys conservatively assumes 
-every two functions can be executed concurrently, and thus can introduce false 
-positives. However, we observe that each kernel module has an initialization 
-phase, after which many functions can be called concurrently by other modules. 
-When performing initialization, the kernel module serially initialized the lock 
-and prepare other data for subsequent operations. And thus functions with the 
-lock initialization and functions called by them tend not to be executed 
-concurrently. Based on this observation, \sys extracts all functions 
-reachable from the function with a lock initialization operation such as {\tt 
-spin\_lock\_init()}, and suppose that this function can not be executed in 
-parallel.
+(T2) \sys uses a {\em lock-usage analysis} to filter out false positives by 
+validating concurrency of the kernel code. After mining locking rules, \sys 
+detects data races by checking whether a given variable access violates the 
+locking rules. To reduce false negatives, \sys conservatively assumes every two 
+functions can execute concurrently, and thus can introduce false positives. 
+However, we observe that each kernel module has an initialization phase, after 
+which many functions can be called concurrently by other modules. When 
+performing initialization, the kernel module serially initializes the locks and 
+prepares other data for subsequent operations. And thus functions with the lock 
+initialization and functions called by them tend not to execute concurrently. 
+Based on this observation, \sys extracts all functions that are reachable from 
+lock initialization functions such as {\tt spin\_lock\_init()} in the function 
+call graph, and suppose that these functions can not execute concurrently.
 
-(T3) \sys uses a {\em pattern-based estimation} to extract harmful races 
-that can trigger memory or logical bugs such as null-pointer dereferences and 
-data inconsistency. Some data races are benign, and developers do not put 
-effort into repairing them because they can not cause serious memory or logical 
-bugs. Therefore, it is important to estimate the harm a data race can bring. We 
-observe that harmful data races can be estimated by given patterns. For 
-example, if a raced data performs as an operand of a dereference instruction, a 
-null-pointer dereference can occur; if a raced data is accessed for more than 
-once and data inconsistency can occur. Based on this observation, \sys 
-uses some patterns to extract harmful races that can cause memory or logical 
-bugs automatically.
+(T3) \sys uses a {\em pattern-based estimation} to extract harmful races that 
+can trigger memory or logic bugs such as null-pointer dereferences, data 
+inconsistencies and double fetches. Some data races are benign, and developers 
+do not put effort into repairing them because they can not cause serious memory 
+or logic bugs. Therefore, it is important to estimate the harmfulness a data 
+race can bring. We observe that harmful data races can be estimated by specific 
+patterns. For example, if a raced data performs as an operand of a dereference 
+instruction, a null-pointer dereference can occur; if a raced data structure is 
+accessed for more than once and each access gets its different fields, a data 
+inconsistency can occur; if a data race is caused by an unprotected write, and 
+it is likely to introduce double fetches.Based on this observation, \sys 
+exploits some patterns to extract harmful data races that can cause memory or 
+logic bugs automatically.
 
@@
 \begin{itemize}
-	\item We analyze the challenges of data race detection in OS kernels, and 
-	propose three key techniques to address	these challenges: (T1) an {\em 
-	alias-aware rule mining method} to automatically deduce locking rules; (T2) 
-	a {\em lock-usage analysis} to filter out false positives caused by 
-	concurrency; (T3) a {\em pattern-based estimation} to extract harmful data 
-	races that can trigger memory or logical bugs such as null-pointer 
-	dereferences and data inconsistency.
+	\item We analyze the challenges to detect races in kernels, and propose 
+	three key techniques to address	these challenges: (T1) an {\em alias-aware 
+	rule mining method} to automatically deduce locking rules; (T2) a {\em 
+	lock-usage analysis} to filter out false positives caused by code that can 
+	not execute concurrently; (T3) a {\em pattern-based estimation} to extract 
+	harmful data races that can cause memory or logic bugs such as null-pointer 
+	dereferences, data inconsistencies and double fetches.		
 	\item Based on these three key techniques, we design a practical static 	
@@
 	estimate the harmfulness of these data races in OS kernels.
-	\item We have evaluated \sys on Linux 6.2, and find xxx data races, 	
-	with a false positive rate of xxx. Among these data races, xxx are 
+	\item We have evaluated \sys on Linux 6.2, and find 273 real data races, 	
+	with a false positive rate of 19.9\%. Among these data races, 88 are 
 	estimated to be harmful. We have reported these harmful bugs to Linux 
-	kernel developers, and xxx of them have been confirmed.
+	kernel developers, and 32 of them have been confirmed.
 \end{itemize}

### 2023-08-02 2.motivation.tex
Finish a temporary version (without comparing experiment and some evaluation results)

@@
 Figure~\ref{fig_demo_bug} shows a real null-pointer dereference caused by a 
-data race in the Linux DRM driver. In the DRM driver, the functions {\tt 
-exynos\_drm\_crtc\_atomic\_disable()} and {\tt exynos\_crtc\_handle\_event()} 
-can execute concurrently. We exploit \textcircled{\footnotesize{n}} to 
-represent the execution order of instructions and show one execution case on 
-the left of Figure~\ref{fig_demo_bug}. In Thread 1, the variables {\tt 
-crtc->state->event} and {\tt crtc->state->active} are checked by an if 
-statement in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. After the 
-condition is calculated to be true, the function {\tt 
-exynos\_crtc\_handle\_event()} is executed in Thread 2. In this function, the 
-value of {\tt crtc->state->active} is assigned to {\tt event} 
+data race in the Linux DRM driver. In this example, We exploit 
+\textcircled{\footnotesize{n}} to represent the execution order of instructions 
+and show one execution case on the left of Figure~\ref{fig_demo_bug}. In the 
+DRM driver, the functions {\tt exynos\_drm\_crtc\_atomic\_disable()} and {\tt 
+exynos\_crtc\_handle\_event()} can execute concurrently. In Thread 1, the 
+variables {\tt crtc->state->event} and {\tt crtc->state->active} are checked by 
+an if statement in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. 
+After the condition is calculated to be true, the function {\tt 
+exynos\_crtc\_handle\_event()} execute in Thread 2. In this function, the value 
+of {\tt crtc->state->event} is assigned to {\tt event} 
 (\textcircled{\footnotesize{2}}) and then {\tt event} is checked in an if 
@@
 (\textcircled{\footnotesize{4}}). Right after this assignment, the function 
-{\tt drm\_crtc\_send\_vblank\_event} is called  in Thread 1 
-(\textcircled{\footnotesize{6}}) with the argument {\tt crtc->state->active}, 
+{\tt drm\_crtc\_send\_vblank\_event()} is called  in Thread 1 
+(\textcircled{\footnotesize{6}}) with the argument {\tt crtc->state->event}, 
 after acquiring the lock {\tt crtc->dev->event\_lock} 
 (\textcircled{\footnotesize{5}}). In the called function, the variable {\tt 
-crtc->state->event} is dereferenced through {\tt e->pipe} 
-(\textcircled{\footnotesize{6}}). In this execution case, the data structure 
-{\tt crtc->state->event} is first assigned with NULL and then dereferenced, and 
-thus a null-pointer dereference can occur.
+crtc->state->event} is finally dereferenced through {\tt e->pipe} 
+(\textcircled{\footnotesize{7}}). In this execution case, the variable {\tt 
+crtc->state->event} is first assigned with a NULL value and then dereferenced, 
+and thus a null-pointer dereference can occur.
 
@@
 
-\PP{C1: Getting locking rules.} The relationship between variables and locks 
-is not well documented in OS kernels, making it hard to determine whether a 
-specific variable should be protected by a lock and which lock is 
-required (locking rules), even for an expert developer. And thus existing 
-annotation-based approaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
+\PP{C1: Getting locking rules.} The relationship between variables and locks is 
+not well documented in OS kernels, making it hard to determine whether a 
+specific variable should be protected by a lock and which lock is required, 
+even for an expert developer. And thus existing 
+annotation-basedapproaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
 Anderson:PLDI09, Zhou:MICRO19, Flanagan:PASTE01, Flanagan:PLDI00, 
@@
 lockset-based analysis to detect data races automatically, but they do not 
-consider alias relations~\cite{Voung:FSE07, Engler:SOSP03} or just use 
+consider alias relationships~\cite{Voung:FSE07, Engler:SOSP03} or just use 
 imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, 	
 Pratikakis:PLDI06, Naik:PLDI06}. However, due to the heavy use of pointers and 
-data structure fields in OS code, the alias relationships between variables can 
-be very complex, and thus lacking effective alias analysis can introduce many 
-false locking rules.
+data structure fields in kernel code, the alias relationships between variables 
+can be very complex, and thus lacking effective alias analysis can introduce 
+both false positives and false negatives.
 
-\PP{C2: Dropping false data races.} Static analysis suffers from false 
-positives. For example, each data race involves more than one code path that 
-should be able to concurrently execute. However, which code can execute 
-concurrently is not well documented for an OS kernel, and it is also hard to 
-determine concurrent code statically due to the complexity of OS code. Thus 
-static analysis can report many false data races.
+\PP{C2: Dropping false data races.} Due to lack of run-time information, static 
+analysis suffers from false positives. For example, each data race involves 
+more than one code path that should be able to concurrently execute. However, 
+which code can execute concurrently is not well documented for an OS kernel, 
+and it is also hard to determine concurrent code statically due to the 
+complexity of OS code. Thus static analysis can report many false data races.
 
-\PP{C3: Estimating harmfulness of data races.} Many data races are benign, and 
-can not cause memory or logic bugs, and thus developers are unwilling to put 
-effort into repairing them. To automatically detect harmful data races, most 
-approaches~\cite{Narayanasamy:PLDI07, Sen:PLDI08, Kasikci:SOSP13, 
-Kasikci:ASPLOS12} estimate the harmfulness of data races through dynamic 
-analysis, and explore thread interleavings to trigger data races to estimate 
-their harmfulness. However, they suffer from low code coverage and thus can 
-miss many real harmful data races.
+\PP{C3: Estimating harmfulness of data races.} Many data races are benign or 
+introduced by developers deliberately to improve the performance of OS kernels, 
+and can not cause memory or logic bugs in fact. Therefore, developers are 
+unwilling to put effort into repairing these data races. To automatically 
+detect harmful data races, most approaches~\cite{Narayanasamy:PLDI07, 
+Sen:PLDI08, Kasikci:SOSP13, Kasikci:ASPLOS12} explore thread inter-leavings 
+dynamically to trigger data races and observe their impact on the checked 
+software. However, they suffer from low code coverage and thus can miss many 
+real harmful data races.

### 2023-08-02 3.technique.tex
Finish a temporary version (without comparing experiment and some evaluation results)

@@
 locking rules. For {\em C2}, we propose a {\em lock-usage analysis} to filter 
-out false data races by validating the concurrency of kernel code. For {\em 
-C3}, we propose a pattern-based estimation to extract harmful races that can 
-trigger memory or logical bugs such as null-pointer dereference and data 
-inconsistency. We introduce them as follows:
+out false positives caused by kernel code that can not execute concurrently. 
+For {\em C3}, we propose a {\em pattern-based estimation} to extract harmful 
+data races that can trigger memory or logic bugs such as null-pointer 
+dereference, data inconsistencies and double fetches. We introduce them as 
+follows:
 
@@
 kernels, but it can be inferred from the kernel code. Specifically, a variable 
-is often protected by the lock stored in the same data structure. And thus if a 
+is often protected by the lock exist in the same data structure. And thus if a 
 variable is accessed after acquiring a lock existing in the same data structure 
@@
 the protecting lock exist in the same data structure can be determined through 
-an alias graph~\cite{Li:ASPLOS22, Kastrinis:CC18} by finding their common 
+an alias graph~\cite{Li:ASPLOS22, Kastrinis:CC18}, by finding their common 
 ancestor. Based on this insight, we propose an {\em alias-aware rule mining 
@@
 precise field-sensitive alias relationships of alias graph, our alias-aware 
-rule mining method can find data structure filed and its protecting lock 
-effectively.
+rule mining method can effectively extract accessed filed and its corresponding 
+lock field.
 
@@
 structure can be determined by finding a common ancestor. If they are in the 
-same data structure, the lock is likely to protect the variable when it is 
-accessed. Figure~\ref{fig_pseudocode_get_access} shows the pseudocode to 
+same data structure, the variable is likely to be protected by the lock. 
+Figure~\ref{fig_pseudocode_get_access} shows the pseudocode to 
 get the field of the accessed variable and the field of the protecting lock in 
@@
 does not exist, the analysis returns a NULL pair (Lines 2-3). Otherwise, the 
-analysis gets the access path for the node of the accessed variable and the 
-node of the lock variable from the common ancestor (Lines 5-7).
+analysis gets the access paths for the node of the accessed variable and the 
+node of the protecting lock from the common ancestor (Lines 5-7).
 
-Take the alias graph in Figure~\ref{fig_demo_alias_graph} as an example, {\tt 
-\&dev->flag} is represented by node $\mathit{n_2}$, and {\tt \&dev->lock} is 
-represented by node $\mathit{n_4}$. The two nodes have a common ancestor 
-$\mathit{n_1}$, and thus the accessed variable {\tt \&dev->flag} and the 
-protecting lock {\tt \&dev->lock} can be inferred to exist in the same data 
-structure (namely {\tt Dev}). Therefore, the structure field Dev.flag is likely 
-to be protected by the lock stored in the structure field Dev.lock.
+Take the alias graph in Figure~\ref{fig_demo_alias_graph} as an example, the 
+accessed variable {\tt dev->flag} is represented by node $\mathit{n_2}$, and 
+the lock {\tt dev->lock} is represented by node $\mathit{n_4}$. The two nodes 
+have a common ancestor $\mathit{n_1}$, and thus the accessed variable {\tt 
+dev->flag} and the lock {\tt dev->lock} can be inferred to exist in the same 
+data structure (namely {\tt Dev}). Therefore, the structure field Dev.flag is 
+likely to be protected by the lock stored in the structure field Dev.lock.
 
@@
 Based on the alias graph, our alias-aware rule mining method performs an 
-inter-procedural path-based~\cite{Li:ASPLOS22}, field-sensitive and alias-aware 
-analysis to effectively mine locking rules about whether a specific variable 
-should be protected by a lock and which lock is required. Overall, our 
-alias-aware rule mining method has three main stages shown in 
-Figure~\ref{fig_workflow}. In Stage 1, it extracts key variables that may be 
-protected by a lock. In Stage 2, it collects all accesses to key fields as well 
-as code paths these accesses exist in. In Stage 3, it deduces locking rules by 
-calculating the proportion of field accesses protected by a specific lock in 
-all field accesses.
+inter-procedural, path-based~\cite{Li:ASPLOS22}, field-sensitive and 
+alias-aware analysis to effectively mine locking rules about whether accesses 
+to a specific data structure field should be protected and which data structure 
+field the protecting lock exist in. Overall, our alias-aware rule mining method 
+has three main stages shown in Figure~\ref{fig_workflow}. In Stage 1, it 
+extracts key fields that may be protected by a lock. In Stage 2, it collects 
+all accesses to key fields as well as code paths these accesses exist in. In 
+Stage 3, it deduces locking rules by calculating the proportion of field 
+accesses protected by a specific lock in all field accesses.
 
-\PP{S1: Key-field extraction.} The OS Kernel has a large code base with 
+\PP{S1: Key-field extraction.} The OS kernel has a large code base with 
 numerous variables. However, only a small part of variables should be protected 
 by a specific lock, and thus collecting all variable accesses can introduce 
-much unnecessary overhead. Generally, only a few field accesses can miss 
-necessary protecting locks, due to the carelessness of developers. Based on 
-this insight, our analysis first extracts key fields that may need to be 
-protected by a specific lock, by performing a lock-set analysis to find whether 
-a given variable is accessed after acquiring a lock in the same data structure 
-as the accessed field in any code path.
+much unnecessary overhead. Generally, given a data structure field, only a 
+small portion of accesses to it can miss necessary protecting locks, due to the 
+carelessness of developers. Based on this insight, our analysis first extracts 
+key fields that may need to be protected by a specific lock, by performing a 
+lock-set analysis to find whether a given variable is accessed after acquiring 
+a lock in the same data structure in any code path.
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{figures/fig_pseudocode_lock_set.pdf}
+	\includegraphics[width=1\linewidth]{figures/fig_pseudocode_lock_set.pdf}
 	\figcaption{Pseudocodes of lock-set analysis.}
@@
 instruction is a lock operation, the node of the operand is inserted into the 
-lock set (Lines 3-4). Otherwise, if the instruction is an unlock operation, the 
-node of the operand is removed from the lock set (Lines 5-6).
+lock set, which stores all nodes of the acquired locks (Lines 3-4). Otherwise, 
+if the instruction is an unlock operation, the node of the operand is removed 
+from the lock set (Lines 5-6).
 
@@
 Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to extract key 
-fields that may be protected by a specific lock, based on lock-set analysis and 
-the alias graph. The analysis starts from each function without a caller 
+fields that may be protected by a specific lock, based on the lock-set analysis 
+and the alias graph. The analysis starts from each function without a caller 
 function (Lines 16-26) and performs a path-based analysis~\cite{Li:ASPLOS22} 
-(Lines 17-25). For each instruction in the code path, the analysis first update 
-the alias graph according to the instruction with the four operations (MOVE, 
-STORE, LOAD and GEP) (Line 1), and then performs a lock-set analysis (Line 2) 
-to get all acquired locks. After updating the alias graph and the lock set, if 
-the instruction is a write or a read, the analysis first gets the node of the 
-operand with the new alias graph (Lines 4-5). And then, for each node of the 
-lock variable in the lock set, the analysis uses the alias graph to extract the 
-protected data structure field (Lines 6-12). If the field is not NULL, it is 
-inserted into the set of key fields (Lines 9-11). For inter-procedural 
-analysis, the analysis copies function definition into its call sites, but this 
+(Lines 17-25). For each instruction in the code path, the analysis first 
+updates the alias graph according to the instruction with the four operations 
+(MOVE, STORE, LOAD and GEP) that can change alias relationships (Line 1), and 
+then performs a lock-set analysis (Line 2) to get all acquired locks. After 
+updating the alias graph and the lock set, if the instruction is a write or a 
+read, the analysis first gets the node of the operand with the new alias graph 
+(Lines 4-5). And then, for each node of the acquired lock in the lock set, the 
+analysis uses the alias graph to extract the protected data structure field 
+(Lines 6-12). If the field is not NULL, it is inserted into the set of key 
+fields (Lines 9-11). Note that to perform inter-procedural analysis, the 
+analysis copies each function definition into its call sites, but this 
 operation is not shown in Figure~\ref{fig_pseudocode_field_extract} for 
@@
 \PP{S2: Field-access collection.} With the key fields extracted in Stage 1, the 
-analysis only needs to collect accesses to the variables that are protected in 
-some cases, because if a variable is never protected by any lock when is 
-accessed, it is less likely to be shared by different threads, and thus can not 
-introduce any concurrency issue.
+analysis only needs to collect accesses to the data structure field that are 
+protected in some cases, because if a field is never protected by any lock when 
+is accessed, it is less likely to be shared by different threads, and thus can 
+not introduce any concurrency issue.
 
@@
 	\includegraphics[width=1\linewidth]{figures/fig_pseudocode_access_collect.pdf}
-	\figcaption{Pseudocodes of key-field extraction.}
+	\figcaption{Pseudocodes of field-access collection.}
 	\label{fig_pseudocode_access_collect}
@@
 to the handled instruction (Lines 1-2). And then, the access type (either a 
-write or a read) is set according to the instruction (Lines 5-8). However, if 
-the instruction is not an access operation, the function returns a NULL value 
-for the field access (Line 10). Otherwise, the analysis first gets the node of 
-the operand with the new alias graph (Lines 12-13), and then for each node of 
-the lock variable in the lock set, the analysis uses the alias graph to extract 
-the fields that the accessed variable and the lock variable stored in (Lines 
-16-17). If the fields are found, the field access is returned with the found 
-fields (Lines 18-22). If the fields are not found, the accessed variable  are 
+write or a read) of a field access is set according to the instruction (Lines 
+5-8). However, if the instruction is not an access operation, the function 
+returns a NULL value for the field access (Line 10). Otherwise, the analysis 
+first gets the node of the operand with the new alias graph (Lines 12-13), and 
+then for each node of the acquired lock in the lock set, the analysis uses the 
+alias graph to extract the fields that the accessed variable and the acquired 
+lock stored in (Lines 16-17). If the fields are found, the field access is 
+returned with the found fields (Lines 18-22). Otherwise, the accessed field is 
 not protected by any lock. In this case, if the accessed variable is stored in 
-a key field, the field access is returned with a NULL lock variable (Lines 
-25-30).
+a key field, the field access is returned with a NULL lock (Lines 25-30).
 
@@
 different program sites is not fine-grained enough, because the access to a 
-variable and the acquirement to a specific lock are often packaged in a special 
+field and the acquirement to a specific lock are often packaged in a special 
 function, and all accesses under different calling context are regarded as the 
-same access. On the other hand, distinguishing different accesses by different 
-code paths can also suffer from inaccuracy when a function contains many branch 
-statement and cause numerous code paths for the same variable access. Based on 
-this consideration, the analysis distinguishes different accesses to the same 
-data structure field by different calling contexts. Specifically, given a key 
-field {\em f} and a lock {\em l}, the analysis first finds all access to {\em 
-f} with lock {\em l} from the collected field accesses in Stage 2, and gets the 
-number {\em locked\_access\_num} by counting different calling contexts 
-extracted from the code paths of these field accesses. Then, the analysis finds 
-all access to {\em f} (no matter which lock is accessed), and gets the number 
-{\em all\_access\_num} in the same way as {\em locked\_access\_num}. If the 
-ratio {\em locked\_access\_num}/{\em all\_access\_num} is larger than a given 
+same access in this strategy. On the other hand, distinguishing accesses by 
+different code paths can also suffer from inaccuracy when a function contains 
+many branch statements, because accesses to the same field is regarded as 
+different in different code paths, causing numerous accesses to the same field 
+due to large quantities of code paths. Based on this consideration, the 
+analysis distinguishes accesses to the same data structure field by different 
+calling contexts. Specifically, given a key field {\em f} and a lock {\em l}, 
+the analysis first finds all access to {\em f} with lock {\em l} from the 
+collected field accesses in Stage 2, and gets the number {\em 
+locked\_access\_num} by counting different calling contexts extracted from the 
+code paths of these field accesses. Then, the analysis finds all access to {\em 
+f} (no matter whether a lock is held), and gets the number {\em 
+all\_access\_num} in the same way as {\em locked\_access\_num}. If the ratio 
+{\em locked\_access\_num} / {\em all\_access\_num} is larger than a given 
 threshold, and there is at least one write access to {\em f}, the data 
@@
 \noindent{\textbf{\em Example.}} We use an example in 
-Figure~\ref{fig_demo_rule_mining} to illustrate how to mine locking rules 
-with the three stages. The analysis first performs a path-based analysis to 
-extract key fields. Take the code path {\tt Line 12, Line 6, Line 7, Line 8, 
-Line 9} as an example. The analysis updates the alias graph by handling two GEP 
-operations ({\em \&dev->lock} and {\em fp = dev->flag}) and a STORE operation 
-({\em *fp = f}). The final alias graph is shown in 
-Figure~\ref{fig_demo_rule_mining}(b). At the same time as alias analysis, the 
-analysis records the acquired lock {\em Dev.lock} into the lock set after the 
-lock instruction at Line 6. When analyzing the read instruction at Line 7, the 
-analysis finds the node of the accessed field ($\mathit{n_2}$) and the node of 
-the lock stored in the lock set ($\mathit{n_4}$) has the same ancestor 
-($\mathit{n_1}$), and thus {\em Dev.flag} is a key field. Similarly, {\em 
-Dev.flag*} is also a key field, due to the write instruction at Line 8.
+Figure~\ref{fig_demo_rule_mining} to illustrate how to mine locking rules with 
+the three stages. The analysis first performs a path-based analysis to extract 
+key fields. Take the code path {\tt Line11, Line 12, Line 5, Line 6, Line 7, 
+Line 8, Line 9} as an example. The analysis first updates the alias graph by 
+handling two GEP operations ({\em \&dev->lock} and {\em fp = dev->flag}) and a 
+STORE operation ({\em *fp = f}), and the final alias graph is shown in 
+Figure~\ref{fig_demo_rule_mining}(b). And then at the same time as alias 
+analysis, the analysis records the acquired lock {\em Dev.lock} into the lock 
+set after the lock instruction at Line 6. When analyzing the read instruction 
+at Line 7, the analysis finds the node of the accessed field ($\mathit{n_2}$) 
+and the node of the lock stored in the lock set ($\mathit{n_4}$) has the same 
+ancestor ($\mathit{n_1}$), and thus {\em Dev.flag} is a key field. Similarly, 
+{\em Dev.flag*} is also a key field, due to the write instruction at Line 8.
 
 After extracting key fields, the method collects all accesses to key fields 
-with lock set analysis and the alias graph. Take the code path {\tt Line 15, 
-Line 6, Line 7, Line 8, Line 9, Line 16 and Line 17} as an example, through a 
-lock instruction at Line 6, the analysis records the acquired lock {\em 
-Dev.lock} into the lock set. When analyzing the read instruction at Line 7, the 
-analysis finds the node of the accessed field ($\mathit{n_2}$) and the node of 
-the lock stored in the lock set ($\mathit{n_4}$) has the same ancestor 
-($\mathit{n_1}$), and thus records a field access <{\em Dev.flag}, {\em 
-Dev.lock}>. Similar to the read instruction at Line 7, the method also records 
-a field access <{\em Dev.flag*}, {\em Dev.lock}> for the write instruction at 
-Line 8. Then through an unlock operation at Line 9, the analysis removes the 
-lock {\em Dev.lock} from the lock set. As a result, the key fields {\em 
-Dev.flag} and {\em Dev.flag*} are accessed without acquiring any lock, and thus 
-the analysis records two field accesses <{\em Dev.flag}, {\em NULL}> and <{\em 
-Dev.flag*}, {\em NULL}>. The final field accesses are shown in 
+with lock-set analysis and the alias graph. Take the code path {\tt Line 14, 
+Line 15, Line 5, Line 6, Line 7, Line 8, Line 9, Line 16 and Line 17} as an 
+example, through a lock instruction at Line 6, the analysis records the 
+acquired lock {\em Dev.lock} into the lock set. When analyzing the read 
+instruction at Line 7, the analysis finds the node of the accessed field 
+($\mathit{n_2}$) and the node of the lock stored in the lock set 
+($\mathit{n_4}$) has the same ancestor ($\mathit{n_1}$), and thus records a 
+field access <{\em Dev.flag}, {\em Dev.lock}>. Similar to the read instruction 
+at Line 7, the method also records a field access <{\em Dev.flag*}, {\em 
+Dev.lock}> for the write instruction at Line 8. Then through an unlock 
+operation at Line 9, the analysis removes the lock {\em Dev.lock} from the lock 
+set. As a result, the key fields {\em Dev.flag} and {\em Dev.flag*} are 
+accessed without acquiring any lock at Lines 16 and 17, and thus the analysis 
+records two field accesses <{\em Dev.flag}, {\em NULL}> and <{\em Dev.flag*}, 
+{\em NULL}>. The final field accesses are shown in 
 Figure~\ref{fig_demo_rule_mining}(c).
@@
 Dev.flag} and {\em Dev.flag*} are accessed under three calling contexts, and 
-two of them are protected by the lock {\em Dev.lock}. And thus the two fields 
-{\em Dev.flag} and {\em Dev.flag*} are deduced to be protected by the lock {\em 
-Dev.lock} (assume the threshold of the ratio {\em locked\_access\_num} / 
-{\em all\_access\_num} is 0.6). However, the accesses at Lines 16 and 17 are 
-not protected by {\em Dev.lock} and thus causing two data races.
+two of them are protected by the lock {\em Dev.lock}. Besides, there is a write 
+to {\em Dev.flag*} at Line 8, and thus the field {\em Dev.flag*} is deduced to 
+be protected by the lock {\em Dev.lock} (assume the threshold of the ratio {\em 
+locked\_access\_num} / {\em all\_access\_num} is 0.6). However, the access to 
+{\em Dev.flag*} at Line 17 is not protected by {\em Dev.lock}, causing a data 
+race.
 
@@
 \label{subsec_lock_usage_analysis}
-Our analysis takes functions that have no caller function as the entry as 
-existing work~\cite{Li:ASPLOS22}, and performs a path-based analysis. It 
-assumes that all codes reaching from entry functions can execute concurrently 
-to reduce false negatives. However, this assumption is too conservative and can 
-introduce many false positives because not all entry functions can execute 
-concurrently in fact. We observe that each kernel module has an initialization 
-phase, after which many functions can be called concurrently by other modules. 
-When performing initialization, the kernel module serially initialized the lock 
-and prepare other data for subsequent operations. And thus functions with the 
-lock initialization and functions called by them tend not to be executed 
-concurrently. Based on the observation, we propose a lock-usage analysis to 
-filter out false positives caused by code that can not execute concurrently.
+Our analysis takes functions that have no caller function as the entries like 
+existing work~\cite{Li:ASPLOS22}, and performs a path-based analysis starting 
+from these entry functions. It assumes that all codes reaching from entry 
+functions can execute concurrently, to reduce false negatives. However, this 
+assumption is too conservative and can introduce many false positives because 
+not all entry functions can execute concurrently in fact. We observe that each 
+kernel module has an initialization phase, after which many functions can be 
+called concurrently by other modules. When performing initialization, the 
+kernel module serially initializes the locks and prepares other data for 
+subsequent operations. And thus functions with the lock initialization and 
+functions called by them tend not to execute concurrently. Based on this 
+observation, we propose a lock-usage analysis to filter out false positives 
+caused by code that can not execute concurrently.
 
@@
 Our lock-usage analysis uses the union-find set~\cite{Galler:ACM64} to get all 
-functions that are reachable from the lock initialization function such as {\em 
-mutex\_init()} and {\em spin\_lock\_init()}. 
+functions that are reachable from the lock initialization functions such as 
+{\em spin\_lock\_init()} in the function call graph. 
 Figure~\ref{fig_pseudocode_lock_usage} shows the pseudocodes of the lock-usage 
@@
 Many data races are benign or introduced by developers deliberately to improve 
-kernel performance. They can not cause memory or logic bugs, and thus 
-developers are unwilling to put effort into repairing them. We observe that 
-harmful data races match some typical patterns, and propose a {\em 
+the performance of OS kernels. They can not cause memory or logic bugs in fact, 
+and thus developers are unwilling to put effort into repairing them. We observe 
+that harmful data races match some typical patterns, and propose a {\em 
 pattern-based estimation} to extract data races that can cause memory or logic 
@@
 At present, we propose three patterns that can cause null-pointer dereference, 
-data inconsistency and unprotected write, because these three patterns are 
+data inconsistencies and double fetches, because these three patterns are 
 common and dangerous in OSes. The three patterns are shown in 
@@
 \begin{itemize}
-	\item \PP{Null-pointer dereference.} The two accesses to {\em dev->event} 
+	\item \PP{P1: Null-pointer dereference.} The two accesses to {\em 
+	dev->event} 
 	are not protected by {\em dev->lock}. This can cause a null-pointer 
@@
 	and then dereferenced, a possible null-pointer dereference can occur.
-	%\item \PP{Infinite loop.} The condition of the while statement does not 
-	%protected by {\em dev->lock}. This can cause a infinite loop when {\em 
-	%dev->cnt} is assigned with a non-zero value every time before it is 
-	%checked 
-	%by the while statement. The analysis detects such bugs by judge whether 
-	%the 
-	%data race acts as a condition of a loop statement.
-	\item \PP{Data inconsistency.} Two different fields of the same data 
+	\item \PP{P2: Data inconsistency.} Two different fields of the same data 
 	structure are accessed without acquiring {\em dev->lock}. This can cause a 
@@
 	lock.
-	\item \PP{Unprotected write.} The write access to {\em dev->event} is not 
-	protected by {\em dev->lock}. This is dangerous because the value {\em 
-	dev->event} can be modified at any time when other threads access it, and 
-	thus introduce unpredictable behavior. The analysis detects such a pattern 
-	by judging whether a data race occurs in a write access.
+	\item \PP{P3: Double fetch.} The write access to {\em dev->event} is not 	
+	protected by {\em dev->lock}. This is dangerous because the value of {\em 	
+	dev->event} can be modified at any time when other threads access it. If it 
+	is changed during two accesses in another thread, a double fetch can occur, 
+	introducing unpredictable behavior. The analysis detects such a pattern by 
+	judging whether a data race occurs in a write access.
 \end{itemize}

### 2023-08-02 4.framework.tex
Finish a temporary version (without comparing experiment and some evaluation results)

@@
 estimate the harmfulness of these data races. We have implemented \sys with 
-Clang~\cite{clang} and Z3~\cite{z3}. \sys automatically mines locking rules, 
+Clang~\cite{clang} and Z3~\cite{z3}. \sys can automatically mine locking rules, 
 detects data races and estimates the harmfulness of detected data races. 
@@
 \PP{P2: Locking-rule Mining.} The locking-rule miner uses our alias-aware rule 
-mining method to deduce locking rules about whether a variable access should be 
-protected by a lock and which lock is required.
+mining method to automatically deduce locking rules about whether accesses to a 
+specific data structure field should be protected and which data structure 
+field the protecting lock exist in
 

### 2023-08-02 5.evaluation.tex
Finish a temporary version (without comparing experiment and some evaluation results)

@@
 				\hline
-				\multirow{3}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
+				\multirow{2}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
 				Key fields / total variables& 
@@
 				& Mined locking rules &
-				\\ \cline{2-3}
-				& Protected field accesses / all accesses & 0.6
 				\\ \cline{2-3}
@@
 				\hline
-				\multirow{4}{1.5cm}{\textbf{{\em Data race estimation}}}
+				\multirow{4}{1.5cm}{\textbf{{\em Harmfulness estimation}}}
 				& Null-pointer dereference (confirmed / all) & 15 / 20
 				\\ \cline{2-3}
-				%& Infinite loop (confirmed / all) & 0 / 1
-				%\\ \cline{2-3}
 				& Data inconsistency (confirmed / all) & 7 / 10
 				\\ \cline{2-3}
-				& Unprotected write (confirmed / all) & 10 / 57
+				& Double fetch (confirmed / all) & 10 / 57
 				\\ \cline{2-3}
@@
 We configure \sys with common lock-acquiring/release functions (like {\tt 
-spin\_lock and spin\_unlock}) to perform lock-set analysis to extract key 
-fields and mine locking rules, and lock-initialization functions (like {\tt 
-spin\_lock\_init}) to filter our false data races caused by code paths that 
-can not execute concurrently. And then run \sys to automatically check the 
-kernel source code. We manually check all the data races found by \sys, and 
+spin\_lock} and {\tt spin\_unlock}) to perform lock-set analysis to extract key 
+fields and collect field accesses, and lock-initialization functions (like {\tt 
+spin\_lock\_init}) to filter our false data races caused by code that can not 
+execute concurrently. And then run \sys to automatically check the kernel 
+source code. We manually check all the data races found by \sys, and 
 Table~\ref{tbl_bug_detection} shows the results, and source code lines are 
@@
 variables. Handling all variables when mining locking rules can introduce much 
-overhead. However, we observe that a variable tends to be protected by the lock 
-stored in the same data structure as the accessed variable. Based on this 
-observation, our locking-rule mining method first extracts a key field by 
-finding whether there exists any access to it that is protected by a lock 
-stored in the same data structure. This method drops xxx\% variables (xxx out 
-xxx) that need to be handled when mining locking rules, and thus can reduce 
-overhead significantly. After extracting key fields, our locking-rule mining 
-method collects all accesses to these key fields, and then deduces locking-rule 
-based on statistics. In this paper, given a data structure field, we set the 
+overhead. However, we observe that the variable that is accessed and the lock 
+to protect the access often exist in the same data structure. Based on this 
+observation, our locking-rule mining method extracts key fields by finding 
+whether there exists any access to it that is protected by a lock stored in the 
+same data structure. This method drops xxx\% variables (xxx out xxx) that need 
+to be handled when mining locking rules, and thus can reduce overhead 
+significantly. After extracting key fields, our locking-rule mining method 
+collects all accesses to these key fields, and then deduces locking rules based 
+on statistics. In this paper, given a data structure field, we set the 
 threshold of the ratio of accesses protected by a specific lock to all accesses 
 to 0.6. Our alias-aware rule mining method can drop many false rules and thus 
-only mines xxx rules, which can effectively reduce false data races.
+only mines xxx locking rules, which can effectively reduce false data races.
 
@@
 them are real, with a false positive rate of 19.9\%. Besides, our lock-usage 
-analysis drops 63 false data races and the results show that it can reduce 
-false positives significantly.
+analysis drops 63 false data races and the results show that this strategy can 
+reduce false positives significantly.
 
@@
 them. We exploit three patterns to detect null-pointer dereferences, data 
-inconsistency and unprotected write as introduced in 
+inconsistencies and double fetches as introduced in 
 Section~\ref{subsec_estimation}, and find 88 data races in these patterns. We 
@@
 vectors to represent different data structure fields. However, in the LLVM 
-bytecode, accesses to these vectors are divided into a LOAD operation and 
+bytecode, accesses to these vectors are divided into a load operation and 
 several bit operations, and thus \sys can not distinguish accesses to different 
@@
 Apart from false data races caused by the two main reasons, there are still 7 
-other data races. \sys regards a function without caller functions in a kernel 
-module as the entry function and starts analysis from these entries. However, 
+other data races. \sys regards functions without caller functions in a kernel 
+module as the entry functions and starts analysis from these entries. However, 
 in 4 cases, an assertion is put at the beginning of the entry function, to 
@@
 Finally, \sys exploits a lock-usage analysis to filter out false data races 
-caused by code paths that can not execute concurrently. Specifically, the 
-lock-usage analysis extracts all functions that are reachable from the lock 
-initialization functions in a call graph with the union-find set, and assumes 
-that these functions can not execute concurrently. However, some functions can 
-be called at multiple program sites, and some of the program sites may be not 
-reachable from the lock initialization functions, but data races that exist in 
-these functions are all dropped by our lock-usage analysis.
+caused by code that can not execute concurrently. Specifically, the lock-usage 
+analysis extracts all functions that are reachable from the lock initialization 
+functions in a call graph with the union-find set, and assumes that these 
+functions can not execute concurrently. However, some functions can be called 
+at multiple program sites, and some of the program sites may be not reachable 
+from the lock initialization functions, but data races that exist in these 
+functions are all dropped by our lock-usage analysis.
 
@@
 calculated to be true, the variable {\em codec->regmap} is set to be NULL by 
-the function {\em snd\_hdac\_regmap\_exit()} in Thread 2. And then the function 
-{\em regcache\_sync()} is called in Thread 1, with the argument {\em 
+the function {\em snd\_hdac\_regmap\_exit()} in Thread 1. And then the function 
+{\em regcache\_sync()} is called in Thread 2, with the argument {\em 
 codec->regmap}, after acquiring the lock {\em codec->regmap\_lock}. In the 
-called function, the variable {\em codec->regmap\_lock} is dereferenced through 
+called function, the variable {\em codec->regmap} is dereferenced through 
 {\em map->lock()}. In this execution case, the data structure field {\em 
@@
 
-\PP{Unprotected write in the LPFC SCSI driver.} In 
-Figure~\ref{fig_case_bugs}(c), the value of the shared variable {\em 
-phba->fcf.fcf\_flag} is set to 0 by the function {\em 
-lpfc\_unregister\_fcf\_rescan()}, without acquiring the lock {\em 
-phba->hbalock}, and this can cause double-fetch in several program sites. For 
-example, if {\em phba->fcf.fcf\_flag} is set to 0 by Thread 2 between the two 
-accesses to it in the function {\em lpfc\_register\_fcf()} in Thread 1 (Lines 
-1873 and 1874), the values of the two accesses can be different, causing 
-unpredictable behaviors.
+\PP{Double fetch in the LPFC SCSI driver.} In Figure~\ref{fig_case_bugs}(c), 
+the value of the shared variable {\em phba->fcf.fcf\_flag} is set to 0 by the 
+function {\em lpfc\_unregister\_fcf\_rescan()} in Thread 2, without acquiring 
+the lock {\em phba->hbalock}, and this can cause double fetches in several 
+program sites. For example, if {\em phba->fcf.fcf\_flag} is set to 0 by Thread 
+2 between the two accesses to it in the function {\em lpfc\_register\_fcf()} in 
+Thread 1 (Lines 1873 and 1874), the values of the two accesses can be 
+different, causing unpredictable behaviors.

### 2023-08-02 6.discussion.tex
Finish a temporary version (without comparing experiment and some evaluation results)

@@
 \PP{Benefits in other aspects.} \sys uses an alias-aware mining method to 
-deduce precise locking rules about whether accesses to a variable should be 
-protected by a lock, and which lock is required. Such rules can benefit in 
-various aspects.
+deduce precise locking rules about whether accesses to a specific data 
+structure field should be protected and which data structure field the 
+protecting lock exist in. Such rules can benefit in various aspects.
 
@@
 developers can limit the critical zone protected by a lock/unlock pair to the 
-exact accesses that should be protected, and let other code can execute 
+exact accesses that should be protected, and let other code execute 
 concurrently. And these locking rules can also help improve documents of OS 
@@
 
-\PP{Comparison to other approaches.} We have tried our best to experimentally 
-compare \sys to existing works in data race detection including 
-RELAY~\cite{Voung:FSE07}, WHOOP~\cite{Deligiannis:ASE15} and 
-Goblint~\cite{Vojdani:ASE16}. Unfortunately, they are too old and generate too 
-many errors when compiling and running, and thus we compare \sys to them 
-methodologically.
+%\PP{Comparison to other approaches.} We have tried our best to experimentally 
+%compare \sys to existing works in data race detection including 
+%RELAY~\cite{Voung:FSE07}, WHOOP~\cite{Deligiannis:ASE15} and 
+%Goblint~\cite{Vojdani:ASE16}. Unfortunately, they are too old and generate too 
+%many errors when compiling and running, and thus we compare \sys to them 
+%methodologically.
 
@@
 lock, because the accessed variable and the RCU lock are existing in the 
-identical data structure by our alias-aware rule mining method. Second, \sys 
+identical data structure in our alias-aware rule mining method. Second, \sys 
 does not handle function-pointer calls, and thus it can not build complete call 

### 2023-08-02 7.related.tex
Finish a temporary version (without comparing experiment and some evaluation results)

@@
 races automatically, and can help developers focus on data races that can 
-indeed introduce serious issues.
+indeed introduce serious bugs.
 
@@
 
-However, dynamic analysis suffers from low code coverage, and thus the locking 
-rules inferred through logs can be imprecise. Besides, dynamic analyses need to 
-run OS kernels with instrumentation and are hard to deploy.
+However, dynamic analysis suffers from low code coverage, and the locking rules 
+inferred through execution traces such as logs can be imprecise. Besides, 
+dynamic analyses need to run OS kernels with instrumentation and are hard to 
+deploy.

### 2023-08-02 8.conclusion.tex
Finish a temporary version (without comparing experiment and some evaluation results)

--- 2023-07-14 8.conclusion.tex
@@
 \label{sec_conclusion}
+
+In this paper, we develop a novel static approach named \sys to detect data 
+races in OS kernels by mining locking rules. It consists of three key 
+techniques, including an alias-aware rule mining method to automatically deduce 
+locking rules, a lock-usage analysis to filter out false positives caused by 
+code that can not execute concurrently and a pattern-based estimation to 
+extract harmful data races that can trigger memory or logic bugs such as 
+null-pointer dereferences, data inconsistencies and double fetches. In the 
+evaluation, \sys finds 88 real harmful data races, and 32 of them have been 
+confirmed by the developers.

### 2023-08-02 paper.tex
Finish a temporary version (without comparing experiment and some evaluation results)

@@
 system (OS) is expected to handle concurrency, inevitably introducing many 
-concurrent issues. And data race is one of the most harmful concurrent issues 
-that can trigger memory or logical bugs. A data race is caused by wrongly 
+concurrency issues. And data race is one of the most harmful concurrency issues 
+that can trigger memory or logic bugs. Data races are caused by wrongly 
 accessing shared data in different threads. To ensure correct access to shared 
 data, fine-grained locking mechanisms are introduced. However, it is hard to 
-determine whether a lock should be used for a specific variable (locking rules) 
-even for an expert developer, due to poor documentation and complicated logic 
-of OS code. As a result, OS kernel is prone to data races, because necessary 
-locks may be missed by mistake. Static analysis is a common technique to help 
-improve code quality, but it is quite challenging to detect data races in OS 
-kernels automatically, because of lack of knowledge of locking rules and high 
-complexity of concurrent execution.
+determine whether a lock should be held when accessing a specific variable 
+(locking rules) even for an expert developer, because of poor document and 
+complicated logic of OS code. As a result, OS kernel is prone to data races, 
+because necessary locks may be missed by mistake. Static analysis is a common 
+technique to improve code quality, but it is quite challenging to detect data 
+races in OS kernels automatically, due to lack of knowledge of locking rules 
+and high complexity of concurrent execution.
 
 In this paper, we design a practical static analysis approach named \sys, to 
-effectively detect harmful races that can trigger memory or logical bugs in OS 
-kernels by mining locking rules. \sys first employs an alias-aware rule mining 
-method to automatically deduce locking rules, and detects data races caused by 
-violation of these rules. And then performs a lock-usage analysis to filter out 
-false positives caused by concurrency. At last, \sys extracts harmful data 
-races from all detected data races through a pattern-based estimation. We have 
-evaluated \sys on Linux 6.2, and find xxx data races, with a false positive 
-rate of xxx. Among these data races, xxx are estimated to be harmful. We have 
-reported these harmful bugs to Linux kernel developers, and xxx of them have 
-been confirmed.
+effectively detect data races and estimate their harmfulness in OS kernels by 
+mining locking rules. \sys first employs an alias-aware rule mining method to 
+automatically deduce locking rules, and detects data races caused by violation 
+of these rules. And then performs a lock-usage analysis to filter out false 
+positives caused by code that can not execute concurrently. At last, \sys 
+extracts harmful data races from all detected data races through a 
+pattern-based estimation. We have evaluated \sys on Linux 6.2, and find 273 
+data races, with a false positive rate of 19.9\%. Among these data races, 88 
+are estimated to be harmful. We have reported these harmful data races to Linux 
+kernel developers, and 32 of them have been confirmed.
 

### 2023-08-03 1.introduction.tex
Update

@@
 accessed for more than once and each access gets its different fields, a data 
-inconsistency can occur; if a data race is caused by an unprotected write, and 
-it is likely to introduce double fetches.Based on this observation, \sys 
-exploits some patterns to extract harmful data races that can cause memory or 
-logic bugs automatically.
+inconsistency can occur; if a data race is caused by an unprotected write, it 
+is likely to introduce double fetches.Based on this observation, \sys exploits 
+some patterns to extract harmful data races that can cause memory or logic bugs 
+automatically.
 
-We have implemented \sys with LLVM~\cite{clang} and Z3~\cite{z3}. 
-\sys performs automated static analysis on the LLVM bytecode of the 
-checked OS kernel. Overall, we make three main contributions in this paper:
+We have implemented \sys with LLVM~\cite{clang}. \sys performs automated static 
+analysis on the LLVM bytecode of the checked OS kernel. Overall, we make three 
+main contributions in this paper:
 
@@
 	\item We have evaluated \sys on Linux 6.2, and find 273 real data races, 	
-	with a false positive rate of 19.9\%. Among these data races, 88 are 
-	estimated to be harmful. We have reported these harmful bugs to Linux 
+	with a false positive rate of 19.9\%. Among these data races, 87 are 
+	estimated to be harmful. We have reported these harmful data races to Linux 
 	kernel developers, and 32 of them have been confirmed.

### 2023-08-03 2.motivation.tex
Update

@@
 even for an expert developer. And thus existing 
-annotation-basedapproaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
+annotation-based approaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
 Anderson:PLDI09, Zhou:MICRO19, Flanagan:PASTE01, Flanagan:PLDI00, 

### 2023-08-03 3.technique.tex
Update

@@
 kernels, but it can be inferred from the kernel code. Specifically, a variable 
-is often protected by the lock exist in the same data structure. And thus if a 
+is often protected by the lock exists in the same data structure. And thus if a 
 variable is accessed after acquiring a lock existing in the same data structure 
@@
 to a specific data structure field should be protected and which data structure 
-field the protecting lock exist in. Overall, our alias-aware rule mining method 
-has three main stages shown in Figure~\ref{fig_workflow}. In Stage 1, it 
+field the protecting lock exists in. Overall, our alias-aware rule mining 
+method has three main stages shown in Figure~\ref{fig_workflow}. In Stage 1, it 
 extracts key fields that may be protected by a lock. In Stage 2, it collects 
@@
 the analysis deduces locking rules based on statistics. However, on the one 
-hand, distinguishing different accesses to the same data structure field by 
-different program sites is not fine-grained enough, because the access to a 
-field and the acquirement to a specific lock are often packaged in a special 
-function, and all accesses under different calling context are regarded as the 
-same access in this strategy. On the other hand, distinguishing accesses by 
-different code paths can also suffer from inaccuracy when a function contains 
-many branch statements, because accesses to the same field is regarded as 
-different in different code paths, causing numerous accesses to the same field 
-due to large quantities of code paths. Based on this consideration, the 
-analysis distinguishes accesses to the same data structure field by different 
-calling contexts. Specifically, given a key field {\em f} and a lock {\em l}, 
-the analysis first finds all access to {\em f} with lock {\em l} from the 
-collected field accesses in Stage 2, and gets the number {\em 
-locked\_access\_num} by counting different calling contexts extracted from the 
-code paths of these field accesses. Then, the analysis finds all access to {\em 
-f} (no matter whether a lock is held), and gets the number {\em 
-all\_access\_num} in the same way as {\em locked\_access\_num}. If the ratio 
-{\em locked\_access\_num} / {\em all\_access\_num} is larger than a given 
-threshold, and there is at least one write access to {\em f}, the data 
-structure field {\em f} is inferred to be protected by the lock {\em l}. After 
-deducting locking rules, the analysis detects data races by checking whether a 
-field access validates any locking rules.
+hand, distinguishing accesses to the same data structure field by different 
+program sites is not fine-grained enough, because the access to a field and the 
+acquisition of a specific lock are often packaged in a special function, and 
+all accesses under different calling context are regarded as the same access in 
+this strategy. On the other hand, distinguishing accesses by different code 
+paths can also suffer from inaccuracy when a function contains many branch 
+statements, because accesses to the same field is regarded as different in 
+different code paths, causing numerous accesses to the same field due to large 
+quantities of code paths. Based on this consideration, the analysis 
+distinguishes accesses to the same data structure field by different calling 
+contexts. Specifically, given a key field {\em f} and a lock {\em l}, the 
+analysis first finds all access to {\em f} with lock {\em l} from the collected 
+field accesses in Stage 2, and gets the number {\em locked\_access\_num} by 
+counting different calling contexts extracted from the code paths of these 
+field accesses. Then, the analysis finds all access to {\em f} (no matter 
+whether a lock is held), and gets the number {\em all\_access\_num} in the same 
+way as {\em locked\_access\_num}. If the ratio {\em locked\_access\_num} / {\em 
+all\_access\_num} is larger than a given threshold, and there is at least one 
+write access to {\em f}, the data structure field {\em f} is inferred to be 
+protected by the lock {\em l}. After deducting locking rules, the analysis 
+detects data races by checking whether a field access validates any locking 
+rules.
 
@@
 ($\mathit{n_4}$) has the same ancestor ($\mathit{n_1}$), and thus records a 
-field access <{\em Dev.flag}, {\em Dev.lock}>. Similar to the read instruction 
-at Line 7, the method also records a field access <{\em Dev.flag*}, {\em 
-Dev.lock}> for the write instruction at Line 8. Then through an unlock 
-operation at Line 9, the analysis removes the lock {\em Dev.lock} from the lock 
-set. As a result, the key fields {\em Dev.flag} and {\em Dev.flag*} are 
-accessed without acquiring any lock at Lines 16 and 17, and thus the analysis 
-records two field accesses <{\em Dev.flag}, {\em NULL}> and <{\em Dev.flag*}, 
-{\em NULL}>. The final field accesses are shown in 
+field access <{\em Dev.flag}, {\em Dev.lock}, read>. Similar to the read 
+instruction at Line 7, the method also records a field access <{\em Dev.flag*}, 
+{\em Dev.lock}, write> for the write instruction at Line 8. Then through an 
+unlock operation at Line 9, the analysis removes the lock {\em Dev.lock} from 
+the lock set. As a result, the key fields {\em Dev.flag} and {\em Dev.flag*} 
+are accessed without acquiring any lock at Lines 16 and 17, and thus the 
+analysis records two field accesses <{\em Dev.flag}, {\em NULL}, read> and 
+<{\em Dev.flag*}, {\em NULL}, write>. The final field accesses are shown in 
 Figure~\ref{fig_demo_rule_mining}(c).
@@
 functions can execute concurrently, to reduce false negatives. However, this 
-assumption is too conservative and can introduce many false positives because 
-not all entry functions can execute concurrently in fact. We observe that each 
-kernel module has an initialization phase, after which many functions can be 
-called concurrently by other modules. When performing initialization, the 
-kernel module serially initializes the locks and prepares other data for 
+assumption is too conservative and can inevitably introduce false positives 
+because not all entry functions can execute concurrently in fact. We observe 
+that each kernel module has an initialization phase, after which many functions 
+can be called concurrently by other modules. When performing initialization, 
+the kernel module serially initializes the locks and prepares other data for 
 subsequent operations. And thus functions with the lock initialization and 
@@
 data race filtered out by our lock-usage analysis in the Linux DRM scheduler. 
-For example in {\em drm\_sched\_entity\_set\_priority()}, the accesses to {\em 
+For example in {\em drm\_sched\_entity\_set\_priority()}, the access to {\em 
 entity->priority} is protected by the lock {\em entity->rq\_lock} in most 

### 2023-08-03 4.framework.tex
Update

@@
 Based on the three key techniques in Section~\ref{sec_technique}, we develop a 
-practical static approach named \sys, to detect data races in OS kernels, and 
-estimate the harmfulness of these data races. We have implemented \sys with 
-Clang~\cite{clang} and Z3~\cite{z3}. \sys can automatically mine locking rules, 
-detects data races and estimates the harmfulness of detected data races. 
-Figure~\ref{fig_architecture} shows the architecture of \sys, which has four 
-phases:
+practical static approach named \sys, to detect data races in OS kernels and 
+estimates the harmfulness of detected data races automatically. We have 
+implemented \sys with Clang~\cite{clang}. Figure~\ref{fig_architecture} shows 
+the architecture of \sys, which has four phases:
 
@@
 rules mined by our locking-rule miner. After finding a possible data race, the 
-data race filter uses our locking-usage analysis to determine whether it is a 
+data race filter uses our locking-usage analysis to check whether it is a 
 false positive. Besides, for a given data race, there may be multiple code 
@@
 repeated data races can be reported. To drop repeated data races, for a new 
-possible data race, the data race filter checks whether its problematic 
+possible data race, the filter checks whether its problematic 
 instruction is identical to any existing data race. If so, this possible data 
-race is regarded as repeated and dropped.
+race is dropped.
 
@@
 and reports data races that can cause memory or logic bugs. With these results, 
-developers can focus on those harmful data races.
+developers can focus on data races that can bring harmful impacts.

### 2023-08-03 5.evaluation.tex
Update

@@
 				\multirow{2}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
-				Key fields / total variables& 
+				Variables in key fields / total variables& 
 				\\ \cline{2-3}
@@
 				\\ \cline{2-3}
-				& Total harmful data races (confirmed / all) & 32 / 88
+				& Total harmful data races (confirmed / all) & 32 / 87
 				\\ \cline{2-3}
@@
 The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
-are not enabled by the {\em allyesconfig} for the x86-64 architecture. We 
-believe that \sys can also find more data races in other architectures with 
-proper configuration.
+are not enabled by the {\em allyesconfig} for the x86-64 architecture.
 
@@
 inconsistencies and double fetches as introduced in 
-Section~\ref{subsec_estimation}, and find 88 data races in these patterns. We 
+Section~\ref{subsec_estimation}, and find 87 data races in these patterns. We 
 report them to developers and 32 of them have been confirmed and fixed by them. 
@@
 \PP{False positives.} \sys reports 68 false data races, and through manually 
-checking these false data races, we find that they are introduced for three 
-main reasons:
+checking these false data races, we find that they are introduced for two main 
+reasons:
 
@@
 functions in a call graph with the union-find set, and assumes that these 
-functions can not execute concurrently. However, some functions can be called 
-at multiple program sites, and some of the program sites may be not reachable 
-from the lock initialization functions, but data races that exist in these 
-functions are all dropped by our lock-usage analysis.
+functions can not execute concurrently. However, A function can be called under 
+different calling contexts, and some of the calling contexts do not contain 
+lock initialization operations, but data races that exist in these functions 
+are all dropped by our lock-usage analysis, if any calling context contains a 
+lock initialization operation.
 
@@
 snd\_hdac\_regmap\_sync()}. Right after the condition of the if statement is 
-calculated to be true, the variable {\em codec->regmap} is set to be NULL by 
-the function {\em snd\_hdac\_regmap\_exit()} in Thread 1. And then the function 
+calculated to be true, the variable {\em codec->regmap} is set to NULL by the 
+function {\em snd\_hdac\_regmap\_exit()} in Thread 1. And then the function 
 {\em regcache\_sync()} is called in Thread 2, with the argument {\em 
@@
 the lock {\em phba->hbalock}, and this can cause double fetches in several 
-program sites. For example, if {\em phba->fcf.fcf\_flag} is set to 0 by Thread 
-2 between the two accesses to it in the function {\em lpfc\_register\_fcf()} in 
-Thread 1 (Lines 1873 and 1874), the values of the two accesses can be 
-different, causing unpredictable behaviors.
+program sites, even if other accesses are all protected by a lock. For example, 
+if {\em phba->fcf.fcf\_flag} is set to 0 by Thread 2 between the two accesses 
+to it in the function {\em lpfc\_register\_fcf()} in Thread 1 (Lines 1873 and 
+1874), the values of the two accesses can be different, causing unpredictable 
+behaviors.

### 2023-08-04 5.evaluation.tex
Add some evaluation results

@@
 				\multirow{2}{1.5cm}{\textbf{{\em Code analysis}}} & 
-				Source files (analyzed/all) & xxxK/xxxK
+				Source files (analyzed/all) & 22,537 / 32,328
 				\\ \cline{2-3}
-				& Source code lines (analyzed/all) & xxxM/xxxM 
+				& Source code lines (analyzed/all) & 13,686K / 16,646K 
 				\\ \cline{2-3}
 				\hline
-				\multirow{2}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
-				Variables in key fields / total variables& 
+				\multirow{3}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
+				Extracted key fields & 23,074
 				\\ \cline{2-3}
-				& Mined locking rules &
+				& Handled variables (variables in key fields / total 
+				variables)  & 23.2M / 281.0M
+				\\ \cline{2-3}
+				& Mined locking rules & 2,757
 				\\ \cline{2-3}
@@
 				\hline
-				\multirow{3}{1.5cm}{\textbf{{\em Time usage}}} & 
+				\multirow{4}{1.5cm}{\textbf{{\em Time usage}}} & 
 				Key-field extraction & 
 				\\ \cline{2-3} 
-				& Data-race detection &
+				& Field-access collection & 
 				\\ \cline{2-3}
-				& Total time &
+				& Data-race detection and estimation &
+				\\ \cline{2-3}
+				& Total time & 772m
 				\\ \cline{2-3}
@@
 
-\PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it in 
-total analyzes xxxM lines of code in xxxK source code files within xxx hours. 
-The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
-are not enabled by the {\em allyesconfig} for the x86-64 architecture.
+\PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it 
+totally analyzes 13.7M lines of code in 22.5K source files within 12.9 hours. 
+The remaining 3.0M lines of code in 9.8K source files are not analyzed, as they 
+are not enabled by {\em allyesconfig} for the x86-64 architecture.
 
@@
 whether there exists any access to it that is protected by a lock stored in the 
-same data structure. This method drops xxx\% variables (xxx out xxx) that need 
-to be handled when mining locking rules, and thus can reduce overhead 
-significantly. After extracting key fields, our locking-rule mining method 
-collects all accesses to these key fields, and then deduces locking rules based 
-on statistics. In this paper, given a data structure field, we set the 
-threshold of the ratio of accesses protected by a specific lock to all accesses 
-to 0.6. Our alias-aware rule mining method can drop many false rules and thus 
-only mines xxx locking rules, which can effectively reduce false data races.
+same data structure. This method extracts 23.1K key fields and drops 91.7\% 
+variables (23.2M out 281.0M) that need to be handled when mining locking rules, 
+and thus can reduce overhead significantly. After extracting key fields, our 
+locking-rule mining method collects all accesses to these key fields, and then 
+deduces locking rules based on statistics. In this paper, given a data 
+structure field, we set the threshold of the ratio of accesses protected by a 
+specific lock to all accesses to 0.6. Our alias-aware rule mining method can 
+drop many false rules and thus only mines 2,757 locking rules, which can 
+effectively reduce false data races.
 

### 2024-04-28 response.md
Add reviews and responses

We sincerely thank all the reviewers for insightful comments, and hope our response can help solve main concerns about our paper.

## Review A
### A1: Difference between existing static lockset analysis and LR-Miner
In Section 1, we have analyzed the limitations of existing static lockset analyses (L1--L3 in Page 1), to motivate solution techniques in LR-Miner. Thus, we believe the difference can be found here. To emphasize the difference, we will add a separate paragraph in Section 7.

### A2: Handling field access
In LLVM bytecode, both `.` and `->` are represented as the getelementptr instruction, which is handled by LR-Miner. Thus, we believe that LR-Miner handles all possible ways of field access in source code. We will explain this in Section 4.

### A3: Dropping false positives
We have introduced the related techniques in "False positive dropping" paragraph in Page 9. First, LR-Miner uses Z3 [79] to validate code-path feasibility, by translating LLVM instructions into SMT constraints and computing their satisfiability. Second, for concurrency, LR-Miner drops races in functions whose names have keywords like 'init' and 'remove', because they are expected to have no concurrency during execution. We implemented these techniques by referring to existing approaches like DCUAF [6] and SADA [8]. In the evaluation, 8.9K and 2.2K false races were dropped by these techniques, respectively. We will provide more implementation and evaluation details in Section 4 and Section 5.1.

### A4: Formal definition of function summary
As described in Page 7, "this function summary records the access paths and related instructions for all the lock-acquire/release function calls and variable accesses in the analyzed function". We will add a formal definition of function summary and make it more clear here.

### A5: Limitations of type-based approach
Field-based analysis is a common type-based approach (used by DCUAF [6] and SADA [8]) that differentiates variables using structure type and field name. In Section 5.4, we have experimentally compared LR-Miner to field-based analysis (LR-Miner_FieldBased), and LR-Miner produces better results in Table 4. Indeed, field-based analysis neglects the alias relationships of different structure fields, and thus identifies many false locking rules. We will provide more discussion about possible alternatives and their limitations in Section 6.

## Review B
### B1: Bug types of harmful races
We have described the reasons of selecting the four bug types in Page 8. The main reason is that these bug types are common and harmful in OS kernels, and existing approaches focus on finding these bugs in sequential code not concurrency code. As described in Section 6, we believe that LR-Miner can support the detection of other concurrency issues including concurrency UAF bug. For example, a possible way of identifying concurrency UAF bug is to check whether a racy variable is handled by a memory-free operation. We will provide more details in Section 3.3 and Section 6.

### B2: Detecting races involving no lock mechanism
LR-Miner focuses on detecting lock-related races, and cannot find race bugs involving non-lock concurrency mechanisms like wait queue and completion mechanisms. We will discuss this limitation in Section 6.

### B3: Breakdown of performance results
LR-Miner benefits from function summary usage to achieve good performance. As described in "Efficiency improvement" paragraph in Page 10, many repeated analyses of the same function are avoided in this way, which can largely accelerate code analysis. In the evaluation, we tried LR-Miner without function summary, and its time usage exceeded our timeout (one week, namely 168 hours). We will provide more details in Section 5.4.

## Review C
### C1: Minor comments
Thanks a lot for your appreciation. We will address these issues.

## Review D
### D1: Recognizing of the same structure field
The same structure field in different functions can be recognized by comparing their access paths (expressed by the higher-layer structure type followed by a sequence of field names). We will provide more details in Section 3.1.

### D2: Missing lock relations without common ancestor
It is indeed a limitation of LR-Miner. We will discuss this in Section 6.

### D3: Interleavings in harmful estimation
The value of the racy variable is uncertain due to non-determinism of concurrent execution, and thus the access to it may cause problems. LR-Miner does not fully consider interleavings in harmful estimation, and just validates the concurrency of functions containing the found races, as described in "False positive dropping" paragraph in Page 9. We think missing full consideration of interleavings can indeed introduce some inaccuracy in harmful estimation. We will discuss this limitation in Section 6.

### D4: Outdated works in comparison
We have investigated papers of static race detection in OS kernels, and found that CPALockator is the only public-available tool published in recent five years. As for other recent papers, we found that just a few of them (like Lockpick [Security'23]) involve (but do not focus on) static race detection in OS kernels, but their tools are not publicly available. We will provide methodology comparison to these papers in Section 7.

### D5: Writing issues
We will address these issues and carefully check the whole paper.
